### 2021-05-13 papers/assembly/I/main.tex
Fleshout text and diagrams through sources and sinks

\documentclass[aps,prd,onecolumn,nofootinbib,letterpaper,preprintnumbers,superscriptaddress,eqsecnum]{revtex4}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage[usenames,dvipsnames]{color}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{dsfont}
\usepackage{tikz-cd}
\usepackage{cleveref}
\usepackage{txfonts}
\usepackage{tikz, tikz-cd}
\usepackage{subfigure}
\usepackage[version=4]{mhchem}
\usetikzlibrary{backgrounds,circuits,circuits.ee.IEC,shapes,fit,matrix}

\tikzstyle{simple}=[-,line width=2.000]
\tikzstyle{arrow}=[-,postaction={decorate},decoration={markings,mark=at position .5 with {\arrow{>}}},line width=1.100]
\pgfdeclarelayer{edgelayer}
\pgfdeclarelayer{nodelayer}
\pgfsetlayers{edgelayer,nodelayer,main}

\tikzstyle{none}=[inner sep=0pt]

\definecolor{lblue}{rgb}{0,250,255}

\tikzstyle{site}=[circle, fill=lblue!20, draw=black, scale=1, minimum size=1cm]
\tikzstyle{catalyst}=[circle, fill=lblue!20, draw=red, scale=1, minimum size=1cm]
\tikzstyle{transition}=[rectangle, fill=gray!30, draw=black, scale=1, minimum size=1cm]
\tikzstyle{morphism}=[rectangle, fill=gray!30, draw=black, scale=1, minimum size=1cm]

\tikzstyle{empty}=[circle,fill=none, draw=none]
\tikzstyle{inputdot}=[circle,fill=black,draw=black, scale=.5]
\tikzstyle{dot}=[circle,fill=black,draw=black]
\tikzstyle{bounding}=[circle,dashed, fill=none,draw=black, scale=9.00]
\tikzstyle{simple}=[-,draw=black,line width=1.000]
\tikzstyle{arrow}=[-,draw=black,postaction={decorate},decoration={markings,mark=at position .5 with {\arrow{>}}},line width=1.000]
\tikzstyle{tick}=[-,draw=black,postaction={decorate},decoration={markings,mark=at position .5 with {\draw (0,-0.1) -- (0,0.1);}},line width=1.000]
\tikzstyle{inputarrow}=[->,draw=black, shorten >=.05cm]
\tikzstyle{sitemap}=[->, dashed, draw=red, ultra thick, shorten >=.05cm]
\tikzstyle{transitionmap}=[->, dashed, draw=blue, ultra thick, shorten >=.05cm]

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{theorem}{Theorem}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\newcommand{\upset}[1]{#1\!\!\uparrow}
\newcommand{\downset}[1]{#1\!\!\downarrow}
\newcommand{\N}{\mathbb{N}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\Min}[1]{\min{(#1)}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\powerset}{\mathcal{P}}

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\TODO}[1]{\noindent\red{\textbf{TODO:}~#1}}

\begin{document}

\title{Supplementary Information: Quantifying the pathways to life using assembly spaces}
\author{Stuart M. Marshall}
\author{Douglas G. Moore}
\author{Alastair R. G. Murray}
\author{Sara I. Walker}
\author{Leroy Cronin}

\maketitle

\section{Introduction}

This formalism arose as a means of rigorously describing the ``simplest" way of assembling a given object by combining basic building blocks.
With this in mind, we consider a universe of objects together with transitions representing ways in which objects can be combined, mutated or disassembled to yield other objects.
In this setting, the concrete rules or laws relating the transitions with the objects is largely neglected herein, though they are certainly necessary for initially constructing the space for consideration.
We quickly come to the conclusion that a Petri network is appropriate when we consider this kind of process in general, though with specific modifications to allow the types of questions in which we are interested to be well defined.
As such, the fundamental mathematical structure we define and explore herein, referred to as an assembly space, can be described as a sourceless, sinkless Petri net with a chosen basis.
The details of this description will, of course, be made precise below.
With assembly spaces defined, we are then able to define an assembly index for each object in the space, relative to the chosen basis, which characterizes how directly that object can be assembled.
Further, we prove several theorems relating to method of bounding the assembly index for a given object and algorithms for computing or approximating it.

[This is Sara testing if she can git cool like the rest of the peeps in elife, I will remove on my next commit]

\section{Petri Nets}

\subsection{Free Commutative Monoids}

The first task we are faced with is how to formally represent the substrates and products of some transition.
We can look to introductory Chemistry for inspiration.
When we consider a process like
\begin{equation*}
    \ce{2H_2} + \ce{O_2} \rightarrow \ce{2H_2O}
\end{equation*}
we are expressing the idea that some chemical process (transition) consumes twice as much $\ce{H_2}$ as $\ce{O_2}$ and yields $\ce{H_2O}$.
The apparent multiplication ($\ce{2H_2}$) and addition ($\ce{2H_2} + \ce{O_2}$) are simply formal constructs use to encode the stoichiometry of the reaction.
The addition here is commutative; we could have written $\ce{O_2} + \ce{2H_2}$, and it would have meant precisely the same thing.

We take a similar approach herein, though mathematically explicit. Let $X$ be a set. This may be, for example, a set of types of molecules $\{\ce{H_2}, \ce{O_2}, \ce{H_2O}\}$ or a set of kinds of building materials and products $\{\textrm{wood}, \textrm{glass pane}, \textrm{nail}, \textrm{picture frame}\}$.
We can then construct formal linear combinations of the elements of $X$ where the coefficients are natural numbers, e.g. $\ce{2H_2} + \ce{O_2}$ or $4\textrm{wood} + \textrm{glass pane} + 8\textrm{nail}$, and only finitely many coefficients are nonzero.

The collection of such formal linear combinations is referred to as the \textit{free commutative monoid} on $X$ over the natural number $\N$, and is typically denoted $\N[X]$.
The set $X$, in this context, is referred to as the set of generators of $\N[X]$.
An element $a \in \N[X]$ can be written as
\begin{equation*}
    a = \sum_{x \in X} a_x x
\end{equation*}
with natural number coefficients $a_x$, only finitely many non-zero.
This construction naturally has a monoidal structure, hence the name.
You can add such linear combinations to gether, e.g. $\ce{H_2} + (\ce{H_2} + \ce{O_2}) = \ce{2H_2} + \ce{O_2}$, and there exists an identity element $0$ such that $0 + a = a + 0 = a$ for all $a \in \N[X]$.
It is important to stress at this point that the linear combinations do not need to ``make sense'', at least not yet.
Taking the molecular example, $\ce{H_2}$, $\ce{7H_2} + \ce{O_2}$ and $\ce{H_2} + \ce{H_2O} + \ce{O_2}$ are permissible linear combinations and so they are elements of $\N[X]$.
This does not mean that we will, in practice, consider all of them.
Restrictions on which linear combinations are consider will come into the picture when we introduce transitions.

For some of the definitions we will present in coming sections, it will be useful to denote the subset of generators appearing in a given linear combinations. That is for a given element $a \in \N[X]$, the subset generators with non-zero coefficients in $a$, denoted $\pi : \N[X] \rightarrow \powerset{X}$ where
\begin{equation*}
    \displaystyle \sum_{x \in X} a_x x \mapsto \{ x \in X ~|~ a_x \ne 0 \}.
\end{equation*}
To clarify this point, $\pi(\ce{2H2} + \ce{O_2}) = \{\ce{H_2}, \ce{O_2}\}$.

Before departing from the topic of the free commutative monoid, we want to note that while it may seem that we are being overly formalistic, we gain some power by taking this approach.
Most importantly, we can consider what happens if you ``map'' one set of objects into another.
As a concrete example, you could imagine mapping each of the molecules $\ce{H_2}$, $\ce{O_2}$ and $\ce{H_2O}$ to the number of Hydrogen atoms they contain, i.e. $2$, $0$, and $2$.
\red{I don't like that example very much, but it's good enough for the moment.}
This mapping process can be ``lifted'' into a map on the linear combinations by distributing the map over addition, so that $\ce{H_2} + \ce{O_2}$ maps to $2 + 0 = 2$. More formally, any function between sets $f : X \rightarrow Y$ can be uniquely extended to a function $\N[f] : \N[X] \rightarrow \N[Y]$ such that
\begin{equation*}
    \sum_{x \in X} a_x x \mapsto \sum_{x \in X} a_x f(x).
\end{equation*}
The value of doing this is that it will allow us to coherently define a map between assembly spaces, to be defined, which will then be useful for proving certain theorems regarding bounds on quantities of interest.

\subsection{Petri Nets}

We wish to represent the idea of assembling objects out of ``simpler'' objects, in some sense of the word.
To formalize this idea, we will build off of the concept of a \textit{Petri net}.
In simple terms, a Petri net is a set of sites which we think of as representing types of objects, e.g. molecules, building materials, characters, etc..., and a set of transitions representing the ways in which sites can be combined to yield other sites.
For the sake of formalism, we somewhat elevate the transitions above the sites in that we mostly consider the ``inputs'' and ``outputs'' of transitions, rather the transitions that a given site participates in.
With this in mind, we formally define a Petri net as follows.

\begin{definition}\label{def:petri}
    A \textbf{Petri net} is a $P$ consisting of
    \begin{itemize}
        \item a set of \textit{sites}, $S(P)$
        \item a set of \textit{transitions}, $T(P)$
        \item a pair of maps $\begin{tikzcd}
                T(P)
                \arrow[r, shift left = 1, "\sigma_P"]
                \arrow[r, shift right = 1, "\tau_P", swap]
                &
                \N[S(P)]
        \end{tikzcd}$.
    \end{itemize}
    We call $\sigma_P$ the source map and $\tau_Q$ the target map.
    Where context is unambiguous we may simply write $\sigma$ and $\tau$, but will generally retain the $T(P)$ and $S(P)$ notation.
\end{definition}

The maps $\sigma$ and $\tau$ assign to each transition a formal linear combination of sites.
Picking up our molecular example from the previous section, we can consider the Petri net with sites $S = \{\ce{H_2}, \ce{O_2}, \ce{H_2O}\}$, a single transition $T = \{\alpha\}$, and source and target maps
\begin{equation*}
    \begin{tikzcd}
        \alpha \arrow[r, mapsto, "\sigma"] & 2\ce{H_2} + \ce{O_2}
        \quad\textrm{and}\quad
        \alpha \arrow[r, mapsto, "\tau"] & 2\ce{H_2O}
    \end{tikzcd}
\end{equation*}

There is a natural way of graphically representing a Petri net as a bipartite multigraph:

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (H2) at (0, 1) {$\ce{H_2}$};
                \node [style=site] (O2) at (0, -1) {$\ce{O_2}$};
                \node [style=site] (H2O) at (4, 0) {$\ce{H_2O}$};
                \node [style=transition] (alpha) at (2, 0) {$\alpha$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (H2) to (alpha);
                \draw [style=arrow, bend left=15] (H2) to (alpha);
                \draw [style=arrow] (O2) to (alpha);
                \draw [style=arrow, bend right=15] (alpha) to (H2O);
                \draw [style=arrow, bend left=15] (alpha) to (H2O);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

We read this as ``\textit{transition $\alpha$ can only proceed if at least two $\ce{H_2}$ and one $\ce{O_2}$ are present in the system, and will yield two $\ce{H_2O}$ when it does}''.
In essence, this is a statement about was can possibly happen, not necessarily what will happen; that's more a question of dynamics.

As a more elaborate example, one which we will repeatedly refer to as we expand on the formalism, we imagine a simple ligation type process where in the characters $A$ and $B$ can be combined into strings and $C$ acts something like a \red{``catalyst''} making some transitions possible if and only if $C$ exists.
Consider the Petri $P$ net described by $S(P) = \{ A, B, C, AA, AB, AAAB \}$ and $T(P) = \{ \alpha, \beta, \gamma, \delta \}$ with source and target maps:

\begin{align*}
    \sigma(\alpha) &= 2A           & \tau(\alpha) &= AA         \\
    \sigma(\beta)  &= A + B        & \tau(\beta)  &= AB         \\
    \sigma(\gamma) &= AA + AB + C  & \tau(\gamma) &= AAAB + C   \\
    \sigma(\delta) &= AAAB         & \tau(\delta) &= A + AA + B
\end{align*}

which we graphically represent as

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A) to (tau1);
                \draw [style=arrow, bend left=15] (A) to (tau1);
                \draw [style=arrow, bend right=15] (A) to (tau2);
                \draw [style=arrow] (tau1) to (AA);
                \draw [style=arrow] (B) to (tau2);
                \draw [style=arrow] (tau2) to (AB);
                \draw [style=arrow] (AA) to (tau3);
                \draw [style=arrow] (AB) to (tau3);
                \draw [style=arrow, bend left=30] (C) to (tau3);
                \draw [style=arrow] (tau3) to (AAAB);
                \draw [style=arrow, bend left=30] (tau3) to (C);
                \draw [style=arrow] (AAAB) to (tau4);
                \draw [style=arrow, bend right=30] (tau4) to (A);
                \draw [style=arrow, bend right=30] (tau4) to (AA);
                \draw [style=arrow, bend left=40] (tau4) to (B);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

This particular example exhibits a number of interesting features.
First, there are transitions wherein one of the input sites is also an output site, namely $\gamma$'s dependence on $C$.
Drawing from chemical parlance, this represents a catalytic effect; the transition can only proceed when the site is present\footnote{The addition of another transitition next to $\gamma$ which does not depend on or yield $C$ can model processes that can proceed without $C$ but is accelerated by it.}.
Second, transitions such as $\delta$ can be viewed as a degradation or disassembly process.
This allows reactions to run in both directions.
To elaborate on this point, we can return to the water example and introduce a second transition allowing for the degradation of water back into Hydrogen and Oxygen:

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (H2) at (0, 1) {$\ce{H_2}$};
                \node [style=site] (O2) at (0, -1) {$\ce{O_2}$};
                \node [style=site] (H2O) at (4, 0) {$\ce{H_2O}$};
                \node [style=transition] (alpha) at (2, 0) {$\alpha$};
                \node [style=transition] (beta) at (6, 0) {$\beta$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (H2) to (alpha);
                \draw [style=arrow, bend left=15] (H2) to (alpha);
                \draw [style=arrow] (O2) to (alpha);
                \draw [style=arrow, bend right=15] (alpha) to (H2O);
                \draw [style=arrow, bend left=15] (alpha) to (H2O);
                \draw [style=arrow, bend right=15] (H2O) to (beta);
                \draw [style=arrow, bend left=15] (H2O) to (beta);
                \draw [style=arrow, bend right=25] (beta.north) to (H2);
                \draw [style=arrow, bend right=15] (beta.north) to (H2);
                \draw [style=arrow, bend left=15] (beta.south) to (O2);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

An element of $\N[S(P)]$ is called a \textbf{marking} of the Petri net.
Markings play an important role in modeling dynamical systems with Petri nets, representing the state of the system at some time.
For example, the linear combination $5\ce{H_2} + 2\ce{O_2}$ might be thought to represent a system with $5$ molecules of $\ce{H_2}$ and $2$ molecules of $\ce{O_2}$ and no $\ce{H_2O}$.
From this element, we can see that, at least in principle, the transition $\alpha$ might carried out at most twice.
The question of whether, under the action of some dynamical model, one marking can be transformed or processed into another is referred to as the reachability problem and is remarkably difficult.
While we will not delve into dynamical aspects of Petri nets or the to-be-defined assembly spaces in this work, dynamics will likely be a key element of the formal theory going forward.

It is natural, once a mathematical structure has been defined, to ask how two instantiations of that structure may be mapped to one another.
There is, of course, a well defined notion of a map between Petri nets, termed here a \textit{Petri net morphism}, which maps sites and transitions of one Petri net into another in a way that is compatible and consistent with the source and target maps of both Petri nets.
These Petri net morphisms will play a crucial roll in the establishment of lower bounds for the assembly index in \cref{sec:assembly-index}.

\begin{definition}\label{def:petri-map}
    A \textbf{Petri net morphism} from a Petri net $P$ to a Petri net $Q$, denoted $f : P \rightarrow Q$, consisists of a pair of functions $T(f) : T(P) \rightarrow T(Q)$ and $S(f): S(P) \rightarrow S(Q))$ such that the following diagrams commute:
    \begin{equation*}
        \begin{tikzcd}
            T(P)
            \arrow[r, "\sigma_P"]
            \arrow[d, "T(f)", swap]
            &
            \N[S(P)]
            \arrow[d, "\N \lbrack S(f) \rbrack"]
            \\
            T(Q)
            \arrow[r, "\sigma_Q"]
            &
            \N[S(Q)]
        \end{tikzcd}
        \quad
        \textrm{and}
        \quad
        \begin{tikzcd}
            T(P)
            \arrow[r, "\tau_P"]
            \arrow[d, "T(f)", swap]
            &
            \N[S(P)]
            \arrow[d, "\N \lbrack S(f) \rbrack"]
            \\
            T(Q)
            \arrow[r, "\tau_Q"]
            &
            \N[S(Q)].
        \end{tikzcd}
    \end{equation*}
    When possible, we will abuse notation and write $f(s)$ or $f(t)$ for sites $s \in S(P)$ and transitions $t \in T(P)$, in lieu of the more verbose $S(f)(s)$ and $T(f)(t)$.
\end{definition}

The commutative diagrams in this definition are what we mean by "compatible and consistent".
They essentially encode the idea that asking for a transition's inputs/outputs and then mapping those inputs/outputs into the new Petri network will give you the same result as first mapping the transition and then asking for the inputs/outputs of the image.
To clarify this point, consider the two Petri nets $P$ and $Q$

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [scale=3] (P) at (12, 0) {$P$};
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};

                \node [scale=3] (Q) at (12, -6) {$Q$};
                \node [style=site] (W) at (-4, -6) {$W$};
                \node [style=site] (X) at (0, -6) {$X$};
                \node [style=site] (Y) at (4, -6) {$Y$};
                \node [style=site] (Z) at (8, -6) {$Z$};
                \node [style=transition] (tau5) at (-2, -6) {$\phi$};
                \node [style=transition] (tau6) at (2, -6) {$\chi$};
                \node [style=transition] (tau7) at (6, -6) {$\psi$};
                \node [style=transition] (tau8) at (10, -6) {$\omega$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A) to (tau1);
                \draw [style=arrow, bend left=15] (A) to (tau1);
                \draw [style=arrow, bend right=15] (A) to (tau2);
                \draw [style=arrow] (tau1) to (AA);
                \draw [style=arrow] (B) to (tau2);
                \draw [style=arrow] (tau2) to (AB);
                \draw [style=arrow] (AA) to (tau3);
                \draw [style=arrow] (AB) to (tau3);
                \draw [style=arrow, bend left=30] (C) to (tau3);
                \draw [style=arrow] (tau3) to (AAAB);
                \draw [style=arrow, bend left=30] (tau3) to (C);
                \draw [style=arrow] (AAAB) to (tau4);
                \draw [style=arrow, bend right=30] (tau4) to (A);
                \draw [style=arrow, bend right=30] (tau4) to (AA);
                \draw [style=arrow, bend left=40] (tau4) to (B);

                \draw [style=arrow, bend left=15] (W) to (tau5);
                \draw [style=arrow, bend right=15] (W) to (tau5);
                \draw [style=arrow, bend left=15] (tau5) to (X);
                \draw [style=arrow, bend right=15] (tau5) to (X);
                \draw [style=arrow, bend left=15] (X) to (tau6);
                \draw [style=arrow, bend right=15] (X) to (tau6);
                \draw [style=arrow] (tau6) to (Y);
                \draw [style=arrow, bend left=20] (Y) to (tau7);
                \draw [style=arrow] (Y) to (tau7);
                \draw [style=arrow, bend right=20] (Y) to (tau7);
                \draw [style=arrow] (tau7) to (Z);
                \draw [style=arrow, bend left=50] (tau7.south) to (Y.south);
                \draw [style=arrow] (Z) to (tau8);
                \draw [style=arrow, bend left=30] (tau8) to (X);
                \draw [style=arrow, bend right=30] (tau8) to (X);
                \draw [style=arrow, bend right=30] (tau8) to (Y);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

We can construct Petri net morphism $f: P \rightarrow Q$ such that
\begin{align*}
    S(f)(\alpha) &= S(f)(\beta) = \chi    & T(f)&(A)    = T(f)(B) = X \\
    S(f)(\gamma) &= \psi                  & T(f)&(AA)   = T(f)(AB) = T(f)(C) = Y \\
    S(f)(\delta) &= \omega                & T(f)&(AAAB) = Z.
\end{align*}

Graphically, this might be depicted as two sets of arrows $T(f)$ and $S(f)$ which map transitions to transitions and sites to sites, respectively colored blue and red.
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [scale=3] (P) at (12, 0) {$P$};
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};

                \node [scale=3] (Q) at (12, -6) {$Q$};
                \node [style=site] (W) at (-4, -6) {$W$};
                \node [style=site] (X) at (0, -6) {$X$};
                \node [style=site] (Y) at (4, -6) {$Y$};
                \node [style=site] (Z) at (8, -6) {$Z$};
                \node [style=transition] (tau5) at (-2, -6) {$\phi$};
                \node [style=transition] (tau6) at (2, -6) {$\chi$};
                \node [style=transition] (tau7) at (6, -6) {$\psi$};
                \node [style=transition] (tau8) at (10, -6) {$\omega$};

                \node [text=red, scale=2] (s) at (-2, -2.5) {$S(f)$};
                \node [text=blue, scale=2] (t) at (0, -3.5) {$T(f)$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, draw=gray!80, bend right=15] (A) to (tau1);
                \draw [style=arrow, draw=gray!80, bend left=15] (A) to (tau1);
                \draw [style=arrow, draw=gray!80, bend right=15] (A) to (tau2);
                \draw [style=arrow, draw=gray!80] (tau1) to (AA);
                \draw [style=arrow, draw=gray!80] (B) to (tau2);
                \draw [style=arrow, draw=gray!80] (tau2) to (AB);
                \draw [style=arrow, draw=gray!80] (AA) to (tau3);
                \draw [style=arrow, draw=gray!80] (AB) to (tau3);
                \draw [style=arrow, draw=gray!80, bend left=30] (C) to (tau3);
                \draw [style=arrow, draw=gray!80] (tau3) to (AAAB);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau3) to (C);
                \draw [style=arrow, draw=gray!80] (AAAB) to (tau4);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (A);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (AA);
                \draw [style=arrow, draw=gray!80, bend left=40] (tau4) to (B);

                \draw [style=arrow, draw=gray!80, bend left=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend right=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend left=15] (tau5) to (X);
                \draw [style=arrow, draw=gray!80, bend right=15] (tau5) to (X);
                \draw [style=arrow, draw=gray!80, bend left=15] (X) to (tau6);
                \draw [style=arrow, draw=gray!80, bend right=15] (X) to (tau6);
                \draw [style=arrow, draw=gray!80] (tau6) to (Y);
                \draw [style=arrow, draw=gray!80, bend left=20] (Y) to (tau7);
                \draw [style=arrow, draw=gray!80] (Y) to (tau7);
                \draw [style=arrow, draw=gray!80, bend right=20] (Y) to (tau7);
                \draw [style=arrow, draw=gray!80] (tau7) to (Z);
                \draw [style=arrow, draw=gray!80, bend left=50] (tau7.south) to (Y.south);
                \draw [style=arrow, draw=gray!80] (Z) to (tau8);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau8) to (X);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau8) to (X);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau8) to (Y);

                \draw [style=sitemap, bend right=30] (A) to (X);
                \draw [style=sitemap, bend right=30] (B) to (X);
                \draw [style=sitemap, bend right=30] (AA) to (Y);
                \draw [style=sitemap, bend right=30] (AB) to (Y);
                \draw [style=sitemap, bend right=30] (C) to (Y);
                \draw [style=sitemap, bend right=30] (AAAB) to (Z);
                \draw [style=transitionmap, bend right=30] (tau1) to (tau6);
                \draw [style=transitionmap, bend right=30] (tau2) to (tau6);
                \draw [style=transitionmap, bend right=30] (tau3) to (tau7);
                \draw [style=transitionmap, bend right=30] (tau4) to (tau8);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

It is easy to see by inspection that $f$ makes the diagrams in \cref{def:petri-map} commute; for example,
\begin{equation*}
    \sigma_Q(T(f)(\alpha)) = 2X = \N[S(f)](\sigma_P(\alpha)).
\end{equation*}

With an understanding of how to map Petri nets to one another, we can now turn to structural properties and relationships, starting with the concept of a subnet.
In essense, a Petri subnet is simply a Petri net whose sets of sites and transitions are subsets of another net and whose source and target maps are compatible.

\begin{definition}\label{def:subnet}
    Let $Q$ be a Petri net. A \textbf{Petri subnet} $P$ of $Q$ consists of
    \begin{itemize}
        \item a subset of sites of $Q$, denoted $S(P)$
        \item a subset of transitions of $Q$, denoted $T(P)$
        \item a pair of maps $\begin{tikzcd}
                T(P)
                \arrow[r, shift left = 1, "\sigma_P"]
                \arrow[r, shift right = 1, "\tau_P", swap]
                &
                \N[S(P)]
        \end{tikzcd}$.
    \end{itemize}
    such that $\sigma_P = \sigma_Q|_{T(P)}$ and $\tau_P = \tau_Q|_{T(P)}$.
\end{definition}

For example, each of the following nets are Petri subnets of our ligation net:
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A1) at (0, 2) {$A$};
                \node [style=site] (AA1) at (0, 6) {$AA$};
                \node [style=transition] (tau11) at (0, 4) {$\alpha$};

                \node [style=site] (A2) at (4, 2) {$A$};
                \node [style=site] (B2) at (5.6, 2) {$B$};
                \node [style=site] (AA2) at (4, 6) {$AA$};
                \node [style=site] (AB2) at (5.6, 6) {$AB$};
                \node [style=transition] (tau12) at (4, 4) {$\alpha$};
                \node [style=transition] (tau22) at (5.6, 4) {$\beta$};

                \node [style=site] (A3) at (9.6, 0) {$A$};
                \node [style=site] (B3) at (11.2, 0) {$B$};
                \node [style=site] (C3) at (12, 6) {$C$};
                \node [style=site] (AA3) at (9.6, 4) {$AA$};
                \node [style=site] (AB3) at (11.2, 4) {$AB$};
                \node [style=site] (AAAB3) at (10.4, 8) {$AAAB$};
                \node [style=transition] (tau13) at (9.6, 2) {$\alpha$};
                \node [style=transition] (tau23) at (11.2, 2) {$\beta$};
                \node [style=transition] (tau33) at (10.4, 6) {$\gamma$};

                \node [style=site] (AAAB4) at (16.8, 0) {$AAAB$};
                \node [style=site] (A4) at (16, 4) {$A$};
                \node [style=site] (B4) at (17.6, 4) {$B$};
                \node [style=site] (AA4) at (16, 8) {$AA$};
                \node [style=site] (AB4) at (17.6, 8) {$AB$};
                \node [style=transition] (tau44) at (16.8, 2) {$\delta$};
                \node [style=transition] (tau14) at (16, 6) {$\alpha$};
                \node [style=transition] (tau24) at (17.6, 6) {$\beta$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A1) to (tau11);
                \draw [style=arrow, bend left=15] (A1) to (tau11);
                \draw [style=arrow] (tau11) to (AA1);

                \draw [style=arrow, bend right=15] (A2) to (tau12);
                \draw [style=arrow, bend left=15] (A2) to (tau12);
                \draw [style=arrow, bend right=15] (A2) to (tau22);
                \draw [style=arrow] (B2) to (tau22);
                \draw [style=arrow] (tau12) to (AA2);
                \draw [style=arrow] (tau22) to (AB2);

                \draw [style=arrow, bend right=15] (A3) to (tau13);
                \draw [style=arrow, bend left=15] (A3) to (tau13);
                \draw [style=arrow, bend right=15] (A3) to (tau23);
                \draw [style=arrow] (tau13) to (AA3);
                \draw [style=arrow] (B3) to (tau23);
                \draw [style=arrow] (tau23) to (AB3);
                \draw [style=arrow] (AA3) to (tau33);
                \draw [style=arrow] (AB3) to (tau33);
                \draw [style=arrow, bend left=30] (C3) to (tau33);
                \draw [style=arrow] (tau33) to (AAAB3);
                \draw [style=arrow, bend left=30] (tau33) to (C3);

                \draw [style=arrow] (AAAB4) to (tau44);
                \draw [style=arrow] (tau44) to (A4);
                \draw [style=arrow, bend left=50] (tau44) to (AA4);
                \draw [style=arrow] (tau44) to (B4);
                \draw [style=arrow, bend left=15] (A4) to (tau14);
                \draw [style=arrow, bend right=15] (A4) to (tau14);
                \draw [style=arrow, bend right=15] (A4) to (tau24);
                \draw [style=arrow] (B4) to (tau24);
                \draw [style=arrow] (tau14) to (AA4);
                \draw [style=arrow] (tau24) to (AB4);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

Petri net morphisms and subnets interact well.
In particular, for a Petri net morphism $f : P \rightarrow Q$, the image $f(P)$ in $Q$ is a Petri subnet with source and target maps $\sigma_{f(P)}$ and $\tau_{f(P)}$ defined by restriction of $\sigma_Q$ and $\tau_Q$, respectively.
Similarly, the preimage of a Petri subnet of $Q$ is a Petri subnet of $P$ with source and target maps defined by restriction of $\sigma_P$ and $\tau_P$.
Continuing the example map from our ligation model,
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [scale=3] (P) at (12, 0) {$P$};
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};

                \node [scale=3] (Q) at (12, -6) {$Q$};
                \node [style=site] (W) at (-4, -6) {$W$};
                \node [style=site, fill=green!20, draw=red!50] (X) at (0, -6) {$X$};
                \node [style=site, fill=green!20, draw=red!50] (Y) at (4, -6) {$Y$};
                \node [style=site, fill=green!20, draw=red!50] (Z) at (8, -6) {$Z$};
                \node [style=transition] (tau5) at (-2, -6) {$\phi$};
                \node [style=transition, fill=yellow!20, draw=red!50] (tau6) at (2, -6) {$\chi$};
                \node [style=transition, fill=yellow!20, draw=red!50] (tau7) at (6, -6) {$\psi$};
                \node [style=transition, fill=yellow!20, draw=red!50] (tau8) at (10, -6) {$\omega$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, draw=gray!80, bend right=15] (A) to (tau1);
                \draw [style=arrow, draw=gray!80, bend left=15] (A) to (tau1);
                \draw [style=arrow, draw=gray!80, bend right=15] (A) to (tau2);
                \draw [style=arrow, draw=gray!80] (tau1) to (AA);
                \draw [style=arrow, draw=gray!80] (B) to (tau2);
                \draw [style=arrow, draw=gray!80] (tau2) to (AB);
                \draw [style=arrow, draw=gray!80] (AA) to (tau3);
                \draw [style=arrow, draw=gray!80] (AB) to (tau3);
                \draw [style=arrow, draw=gray!80, bend left=30] (C) to (tau3);
                \draw [style=arrow, draw=gray!80] (tau3) to (AAAB);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau3) to (C);
                \draw [style=arrow, draw=gray!80] (AAAB) to (tau4);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (A);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (AA);
                \draw [style=arrow, draw=gray!80, bend left=40] (tau4) to (B);

                \draw [style=arrow, draw=gray!80, bend left=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend right=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend left=15] (tau5) to (X);
                \draw [style=arrow, draw=gray!80, bend right=15] (tau5) to (X);
                \draw [style=arrow, draw=red!50, bend left=15] (X) to (tau6);
                \draw [style=arrow, draw=red!50, bend right=15] (X) to (tau6);
                \draw [style=arrow, draw=red!50] (tau6) to (Y);
                \draw [style=arrow, draw=red!50, bend left=20] (Y) to (tau7);
                \draw [style=arrow, draw=red!50] (Y) to (tau7);
                \draw [style=arrow, draw=red!50, bend right=20] (Y) to (tau7);
                \draw [style=arrow, draw=red!50] (tau7) to (Z);
                \draw [style=arrow, draw=red!50, bend left=50] (tau7.south) to (Y.south);
                \draw [style=arrow, draw=red!50] (Z) to (tau8);
                \draw [style=arrow, draw=red!50, bend left=30] (tau8) to (X);
                \draw [style=arrow, draw=red!50, bend right=30] (tau8) to (X);
                \draw [style=arrow, draw=red!50, bend right=30] (tau8) to (Y);

                \draw [style=sitemap, bend right=30, shorten >=2mm] (A) to (X);
                \draw [style=sitemap, bend right=30, shorten >=2mm] (B) to (X);
                \draw [style=sitemap, bend right=30, shorten >=2mm] (AA) to (Y);
                \draw [style=sitemap, bend right=30, shorten >=2mm] (AB) to (Y);
                \draw [style=sitemap, bend right=30, shorten >=2mm] (C) to (Y);
                \draw [style=sitemap, bend right=30, shorten >=2mm] (AAAB) to (Z);
                \draw [style=transitionmap, bend right=30, shorten >=2mm] (tau1) to (tau6);
                \draw [style=transitionmap, bend right=30, shorten >=2mm] (tau2) to (tau6);
                \draw [style=transitionmap, bend right=30, shorten >=2mm] (tau3) to (tau7);
                \draw [style=transitionmap, bend right=30, shorten >=2mm] (tau4) to (tau8);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
the image, outlined in light red, is a subnet of $Q$.

When it comes to describing systems like the construction of a string, or the building of a molecule, it is often classically unphysical for objects to be created from nothing or destroyed.
For that reason, we need to be able to identify aspects of a Petri net that might represent exactly such a process: sources and sinks.
\begin{definition}\label{def:source-sink}
    Let $P$ be a Petri net and let $t$ be a transition.
    We say that $t$ is a \textbf{source} if $\sigma(t) = 0$, and is a \textbf{sink} if $\tau(t) = 0$.
    $P$ is said to be \textbf{sourceless} if it has no sources.
    Similarly, $P$ is \textbf{sinkless} if it has no sinks.
\end{definition}

Graphically, 
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=transition] (source) at (0, 0) {source};
                \node (A) at (2, 1.6) {};
                \node (B) at (2, 0.8) {};
                \node (C) at (2, 0) {};
                \node (D) at (2, -0.8) {};
                \node (E) at (2, -1.6) {};

                \node (F) at (3, 1.6) {};
                \node (G) at (3, 0.8) {};
                \node (H) at (3, 0) {};
                \node (I) at (3, -0.8) {};
                \node (J) at (3, -1.6) {};
                \node [style=transition] (sink) at (5, 0) {sink};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend left=15] (source) to (A);
                \draw [style=arrow, bend left=15] (source) to (B);
                \draw [style=arrow] (source) to (C);
                \draw [style=arrow, bend right=15] (source) to (D);
                \draw [style=arrow, bend right=15] (source) to (E);

                \draw [style=arrow, bend left=15] (F) to (sink);
                \draw [style=arrow, bend left=15] (G) to (sink);
                \draw [style=arrow] (H) to (sink);
                \draw [style=arrow, bend right=15] (I) to (sink);
                \draw [style=arrow, bend right=15] (J) to (sink);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
It follows immediately from the definition of a Petri map that the image of a sourceless Petri net is sourceless, otherwise the $\sigma$-diagram would not commute.
Similarly, the image of a sinkless Petri net is sinkless.

\begin{definition}\label{def:pathway}
    If $P$ is a Petri net, a \textbf{pathway} $\gamma = t_n \ldots t_1$ in $P$ of length $n \ge 1$ is a sequence of transitions $t_i \in T$ such that $\tau(t_i) \ne 0$, $\sigma(t_i) \ne 0$, $m_{i} + \tau(t_i) = n_{i+1} + \sigma(t_{i+1})$ with $\sigma(t_{i+1}) \ne m_{i}$ and $\tau(t_{i+1}) \ne n_{i+1}$. We write $t \in \gamma$ to denote that the transition $t$ appears in the pathway $\gamma$.
\end{definition}

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition, fill=green!20] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition, fill=green!20] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A) to (tau1);
                \draw [style=arrow, bend left=15] (A) to (tau1);
                \draw [style=arrow, bend right=15] (A) to (tau2);
                \draw [style=arrow] (tau1) to (AA);
                \draw [style=arrow] (B) to (tau2);
                \draw [style=arrow] (tau2) to (AB);
                \draw [style=arrow] (AA) to (tau3);
                \draw [style=arrow] (AB) to (tau3);
                \draw [style=arrow, bend left=30] (C) to (tau3);
                \draw [style=arrow] (tau3) to (AAAB);
                \draw [style=arrow, bend left=30] (tau3) to (C);
                \draw [style=arrow] (AAAB) to (tau4);
                \draw [style=arrow, bend right=30] (tau4) to (A);
                \draw [style=arrow, bend right=30] (tau4) to (AA);
                \draw [style=arrow, bend left=40] (tau4) to (B);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

\begin{definition}\label{def:pathway-connected}
    Let $P$ be a Petri net and $x,y \in S(P)$.
    We say that $x$ is \textbf{pathway connected to} $y$ if and only if there exists a pathway $\gamma = t_n \ldots t_1$ such that $\sigma(t_1) = x + m$ and $\tau(t_n) = y + n$ for some $m, n \in \N[S(P)]$.
    The set of all pathways in $P$ from $x$ to $y$ is denoted $\Gamma_x^y$.
    We denote the set of all pathways in $P$ starting from $x$ as $\Gamma_x = \bigcup_{y \in S(P)} \Gamma_x^y$, and all pathways terminating at $y$ as $\Gamma^y = \bigcup_{x \in S(P)} \Gamma_x^y$.
\end{definition}

\begin{lemma}\label{lem:image-of-pathways}
    Let $f: P \rightarrow Q$ be a Petri net morphism and let $x, y \in S(P)$ be pathway connected in $P$.
    Then $f(x)$ and $f(y)$ are pathway connected in $f(P)$.
\end{lemma}
\begin{proof}
    Let $\gamma = t_n \ldots t_1$ be a pathway from $x$ to $y$.
    Given that $\sigma_P(t_i) \ne 0$ and $\tau_P(t_i) \ne 0$, we know that $\sigma_{f(P)} \ne 0$ and $\tau_{f(P)} \ne 0$.
    Since $\sigma_P(t_1) = x + m$, $\sigma_{f(P)}(f(x)) = f(x) + m'$ for some $m' \in \N[S(f(P))]$; similarly, $\tau_{f(P)}(f(y)) = f(y) + n'$ for some $n' \in \N[S(f(P))]$.

    Now, consider two transitions $t_i$ and $t_{i+1}$ in $\gamma$.
    We know then that there exist $m, n \in \N[S(P)]$ such that $m + \tau_P(t_i) = n + \sigma_P(t_{i+1})$ with $m \ne \sigma_P(t_{i+1})$ and $n \ne \tau_P(t_i)$.
    We can then apply $\N[S(f)]$ to each side to yield
    \begin{equation*}
        \N[S(f)](m) + \N[S(f)](\tau_P(t_i)) = \N[S(f)](n) + \N[S(f)](\sigma_P(t_{i+1})).
    \end{equation*}
    Per the commutativity relations imposed by the assembly map,
    \begin{equation*}
        \N[S(f)](\tau_P(t_i)) = \tau_{f(P)}(f(t_i)))
        \quad\textrm{and}\quad
        \N[S(f)](\sigma_P(t_{i+1})) = \sigma_{f(P)}(f(t_{i+1})).
    \end{equation*}
    So, there exist an $m', n' \in \N[S(f(P))]$ such that $m + \tau_{f(P)}(f(t_i)) = n + \sigma_{f(P)}(f(t_{i+1}))$, namely $m' = \N[S(f)](m)$ and $n' = \N[S(f)](n)$.
    Finally, we know that $m' \ne \sigma_{f(P)}(f(t_{i+1}))$ as otherwise $m = \sigma_P(t_{i+1})$, and similarly $n' \ne \tau_{f(P)}(f(t_i))$.

    This shows that a pathway exists from $f(x)$ to $f(y)$, so $f(x)$ and $f(y)$ are pathway connected.
\end{proof}

\begin{definition}\label{def:site-completion}
    Let $P$ be a petri net and let $X \subseteq T(P)$.
    The \textbf{site completion} of $X$ in $P$, denoted $X^*$ is the Petri net defined by $T(X^*) = X$, $S(X^*) = \pi(X)$, $\sigma_{X^*} = \sigma_P|_{X^*}$ and $\tau_{X^*} = \tau_P|_{X^*}$.
\end{definition}

\begin{definition}\label{def:upper-lower-subnet}
    Let $P$ be a Petri net and let $X \subseteq S(P)$.
    The \textbf{upper subnet} of $X$ in $P$, denoted $\upset{X}$, consists of transitions $T(\upset{X}) = \bigcup_{x \in X}\{t \in \gamma ~|~ \gamma \in \Gamma_x\}$, sites $S(\upset{X}) = X \cup S(T(\upset{X})^*)$, $\sigma_{X\uparrow} = \sigma_P|_{T(X\uparrow)^*}$ and $\tau_{X\uparrow} = \tau_P|_{T(X\uparrow)^*}$.
    Similarly, the \textbf{lower subnet} of $X$ in $P$, denoted $\downset{X}$, consists of transitions $T(\downset{X}) = \{t \in \gamma ~|~ \gamma \in \Gamma_X\}$, sites $S(\downset{X}) = X \cup S(T(\downset{X})^*)$, $\sigma' = \sigma|_{T(X\downarrow)^*}$ and $\tau = \tau|_{T(X\downarrow)^*}$.
    If $x$ is an element $x \in P$, we write $\upset{x} = \upset{\{x\}}$ and $\downset{x} = \downset{\{x\}}$.
\end{definition}

\begin{definition}\label{def:minimal}
    Let $P$ be a Petri net. A site $y$ is \textbf{minimal} in $P$ if $\Gamma_x^y = \emptyset$ for all sites $y \ne x$ in $P$.
    The set of all minimal sites of $P$ is denoted $\Min{P}$.
\end{definition}

\begin{definition}\label{def:basis}
    Let $P$ be a Petri net and let $\B$ be a non-empty subset of sites of $P$.
    We say that $\B$ is a \textbf{basis} of $P$ if $P = \upset{\B}$ such that if $U \subseteq \B$ with $P = \upset{U}$ then $U = \B$.
\end{definition}

\begin{lemma}
    Let $P$ be a Petri net. For any basis $\B$ of $P$, $\Min{P} \subseteq \B$.
    Furthermore, if $\Min{P}$ is a basis of $P$, then $\Min{P}$ is the unique basis of $P$.
\end{lemma}
\begin{proof}
    Let $x \in \Min{P}$ and $\B$ be a basis of $P$, then either there exists a $y \in \B$ such that $\Gamma_y^x$ is non-empty, or $x \in \B$.
    However, $\Gamma^x$ is empty by definition of $\Min{P}$, so we have that $x \in B$.
    Since the choice of $x$ was arbitrary, every element of $\Min{P}$ is in the basis $\B$ and thus $\min{P} \subseteq \B$.
    Furthermore, suppose that $P = \upset{\Min{P}}$.
    Then $\B = \Min{P}$, otherwise $\B$ would not be a basis, by definition.
    Since $\B$ was chosen arbitrarily, this holds for every choice of $\B$ and as such $\Min{P}$ is a unqiue basis for $P$.
\end{proof}

\begin{lemma}\label{lem:image-basis}
    Let $P$ be a Petri net with basis $\B$ and let $f : A \rightarrow B$ be a Petri net morphism.
    Then $f(\B)$ is a basis for $f(P)$.
\end{lemma}
\begin{proof}
    Certainly $f(\B)$ is nonempty as $\B$ is nonempty, and of course $f(P)$ must be in the upper set of $f(\B)$.
    This follows immediately from \cref{lem:image-of-pathways} since every site in $P$ is pathway connected to a site in $\B$.
    Now let $U \subset f(\B)$ such that $f(P) = \upset{U}$.
    \TODO{Show that $f(\B)$ is minimal.}
\end{proof}

\section{Assembly Spaces, Subspaces and Maps}\label{sec:assembly-index}

\begin{definition}\label{def:assembly-space}
    An \textbf{assembly space} $(A, \B_A)$ is a sourceless and sinkless Petri net with basis $\B_A$.
    Were possible, we will simply write $A$ to denote an assembly space $(A, \B_A)$.
\end{definition}

\begin{definition}\label{def:assembly-subspace}
    Let $A$ and $A'$ be assembly spaces.
    $A'$ is an \textbf{assembly subspace} of $A$ if $A'$ is a Petri subnet of $A$.
    This relationship is denoted $A' \subseteq A$.
\end{definition}

\begin{definition}\label{def:rooted}
    Let $(A', \B_{A'})$ be an assembly subspace of $(A, \B_A)$.
    We say that $A'$ is rooted in $A$ if $\B_{A'} \subseteq \B_A$.
\end{definition}

\begin{lemma}\label{lem:lower-is-rooted}
    Let $A$ be an assembly space and let $x \in S(A)$.
    Then $\downset{x}$ with basis $\B_A \cap S(\downset{x})$ is a rooted assembly subspace of $A$.
\end{lemma}
\begin{proof}
    By construction, $\downset{x}$ is a Petri subnet of $A$.
    Since $A$ is an assembly space, it is sourceless and sinkless; and given that $\downset{x}$ is a site-completion, $\downset{x}$ is also sourceless and sinkless.
    This gives us that $\downset{x}$ is an assembly subspace of $A$.
    We need only show that $\B_{x\downarrow} = \B_A \cap S(\downset{x}) \subseteq \B_A$ is a basis of $\downset{x}$.
    First, $\B_{x\downarrow}$ is necessarily nonempty since $\B_A$ is a basis in $A$.
    Further, we know that every site in $\downset{x}$ is above $\B_A \cap S(\downset{x})$ since, again $\B_A$ is a basis in $A$.
    \TODO{Show that $\B_{x\downarrow}$ is minimal.}
\end{proof}

\begin{definition}\label{def:assembly-map}
    Let $A$ and $B$ be assembly spaces.
    An \textbf{assembly map} is simply a Petri net map $f: A \rightarrow B$ between assembly spaces.
\end{definition}

\begin{theorem}\label{lem:assembly-image-is-subspace}
    If $f : A \rightarrow B$ is an assembly map between assembly spaces $A$ and $B$, then $f(A)$ is an assembly subspace of $B$.
\end{theorem}
\begin{proof}
    This follows directly from the fact that the image of a Petri net morphism is a Petri subnet, and that the image of a sourceless, sinkless Petri net is itself sourceless and sinkless.
\end{proof}

\begin{theorem}\label{thm:rooted-image}
    Let $A$ be an assembly space, $Z \subseteq A$ a rooted assembly subspace, and $f : A \rightarrow B$ an assembly map.
    Then $(f(Z), S(f)(\B_Z))$ is rooted in $(f(A), S(f)(\B_A)$.
\end{theorem}
\begin{proof}
    First, $f(Z)$ is certainly an assembly subspace of $f(A)$ as we can view $f|_Z$ as an assembly map, and thus \cref{lem:assembly-image-is-subspace} applies.
    To see that $f(Z)$ is rooted, we first note that $S(f)(\B_Z)$ and $S(f)(\B_A)$ are bases for $f(Z)$ and $f(A)$, respectively.
    Further, since $Z$ is rooted in $A$, then $\B_Z \subset \B_A$, so $f(\B_Z) \subseteq f(\B_A)$.
\end{proof}

\section{The Assembly Index}

\begin{definition}\label{def:size}
    Let $A$ be an assembly space.
    The \textbf{size} of $A$ is the cardinality of $T(A)$, denoted $|A|$.
\end{definition}

\begin{definition}\label{def:finiteness}
    Let $A$ be an assembly space and let $x$ be a site in $A$.
    We say that $x$ is \textbf{finite} in $A$ if $\downset{x}$ in $A$ has finite size.
\end{definition}

\begin{definition}\label{def:index}
    Let $A$ be an assembly space and let $x$ be a finite site in $A$.
    The \textbf{assembly index} of $x$ is the minimal size of all rooted assembly subspaces of $A$ containing $x$, and is denoted $c_A(x)$.
    This can be written $c(x)$ when the relevant assembly space is clear from context.
\end{definition}

\subsection{Bounds on the Assembly Index}

\begin{theorem}\label{cor:subspace-index-upper-bound}
    Let $(A, \B_A)$ be an assembly space and let $(B, \B_A)$ be a rooted assembly subspace of $A$.
    For every finite $b \in S(B)$, the assembly index of $b$ in $B$ is at least the assembly index of $b$ in $A$.
    That is, $c_B(b) \ge c_A(b)$ for all $b \in S(B)$
\end{theorem}
\begin{proof}
    We need only show that $c_A(b)$ cannot be greater than $c_B(b)$ for any $b \in S(B)$.
    To see this, let $Z \subseteq B$ be a rooted assembly subspace containing $b$ with $|Z| = c_B(b)$.
    Then $Z$ is also a rooted assembly subspace of $A$ containing $b$, and so $c_A(b)$ can be at most $|Z|$.
    That is, $c_A(b) \le |Z| = c_B(b)$.
\end{proof}

\begin{theorem}\label{thm:image-lower-bound}
    If $f: A \rightarrow B$ is an assembly map, then $c_{f(A)}(f(x)) \le c_A(x)$ for all finite $x \in S(A)$.
    Furthermore, if $f$ is an embedding, then $c_{f(A)}(f(x)) = c_A(x)$.
\end{theorem}
\begin{proof}
    Suppose $x \in S(A)$ is finite, and $Z \subseteq A$ is a rooted with $s \in S(Z)$ and $|Z| = c_A(x)$.
    Restricting $f$ to $Z$ provides an assembly map $f^* : Z \rightarrow B$ such that $f(Z)$ contains $f(x)$, and is rooted in $B$.
    As such, $c_{f(A)}(f(x))$ can be no larger than $|f(Z)|$, so $c_{f(A)}(f(x)) \le c_A(x)$.

    Furthermore, suppose that $f$ is an embedding, and consider any $Z' \subseteq f(A)$ rooted and containing $f(x)$.
    The preimage of $Z'$ under $f$ is then a rooted assembly subspace of $A$ containing $x$, so $c_A(x)$ can be no larger than $|f^{-1}(Z)|$.
    Together with the previous result, we then have $c_{f(A)}(f(x)) = c_A(x)$.
\end{proof}

\begin{corollary}\label{cor:lowerset-index}
    Let $A$ be an assembly space and $x \in S(A)$ be finite.
    Then $c_A(x) = c_{x\downarrow}(x)$.
\end{corollary}
\begin{proof}
    There is a natural embedding $i : \downset{x} \rightarrow A$, so $c_A(x) = c_{x\downarrow}(x)$ for all finite $x \in S(A)$, as per \cref{thm:image-lower-bound}.
\end{proof}

\subsection{Computability and Algorithms}

\begin{theorem}\label{thm:computable}
    If $A$ is an assembly space and $x \in S(A)$ is finite, then $c(x)$ is computable.
\end{theorem}
\begin{proof}
    Every rooted assembly subspace with minimal size and containing $x$ is a minimal, rooted assembly subspace of $\downset{x}$.
    Since $x$ is finite, $\downset{x}$ is finite; that is $\downset{x}$ has finitely many transitions.
    This, together with the fact that $\pi(\sigma(t))$ and $\pi(\tau(t))$ are finite for all $t \in T(A)$, ensures that there areonly finitely many assembly subspaces of $\downset{x}$.
    As such, the set of all rooted assembly subspaces is computable; simply enumerate them.
    The size of each subspace is, of course, computable as one need simply count the transitions in the subspace.
    Finally, finding the minimum value in a finite set of natural numbers is computable.
    Thus, $c(x)$ is computable.
\end{proof}

\TODO{Typeset split-branch algorithm}

\section{Assembly Examples}

\TODO{Come up with a few fun examples.}

\end{document}