### 2021-05-21 papers/assembly/I/main.tex
Annotate some painpoints regarding bases.
Petri net morphisms do not, in general, preserve bases. Given that's
case, we may have to losen our definition of an Assembly space to "a
special Petri net with chosen generating set".

\documentclass[aps,prd,onecolumn,nofootinbib,letterpaper,preprintnumbers,superscriptaddress,eqsecnum]{revtex4}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{dsfont}
\usepackage{tikz-cd}
\usepackage{cleveref}
\usepackage{txfonts}
\usepackage{tikz, tikz-cd}
\usepackage{subfigure}
\usepackage[version=4]{mhchem}
\usetikzlibrary{backgrounds,circuits,circuits.ee.IEC,shapes,fit,matrix}

\tikzstyle{arrow}=[-,postaction={decorate},decoration={markings,mark=at position .5 with {\arrow{>}}},line width=1.100]
\pgfdeclarelayer{edgelayer}
\pgfdeclarelayer{nodelayer}
\pgfsetlayers{edgelayer,nodelayer,main}

\tikzstyle{none}=[inner sep=0pt]

\definecolor{lblue}{rgb}{0,250,255}

\tikzstyle{site}=[circle, fill=lblue!20, draw=black, scale=1, minimum size=1cm]
\tikzstyle{catalyst}=[circle, fill=lblue!20, draw=red, scale=1, minimum size=1cm]
\tikzstyle{transition}=[rectangle, fill=gray!30, draw=black, scale=1, minimum size=1cm]
\tikzstyle{morphism}=[rectangle, fill=gray!30, draw=black, scale=1, minimum size=1cm]

\tikzstyle{empty}=[circle, fill=none, draw=none]
\tikzstyle{inputdot}=[circle, fill=black, draw=black, scale=.5]
\tikzstyle{dot}=[circle, fill=black, draw=black]
\tikzstyle{bounding}=[circle, dashed, fill=none, draw=black, scale=9.00]
\tikzstyle{simple}=[->, draw=black, line width=1]
\tikzstyle{arrow}=[-, draw=black, postaction={decorate}, decoration={markings,mark=at position .5 with {\arrow{>}}}, line width=1.000]
\tikzstyle{tick}=[-, draw=black, postaction={decorate}, decoration={markings,mark=at position .5 with {\draw (0,-0.1) -- (0,0.1);}}, line width=1.000]
\tikzstyle{inputarrow}=[->, draw=black, shorten >=.05cm]
\tikzstyle{sitemap}=[->, dashed, draw=red, ultra thick, shorten >=.05cm]
\tikzstyle{transitionmap}=[->, dashed, draw=blue, ultra thick, shorten >=.05cm]

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{theorem}{Theorem}

\theoremstyle{definition}
\newtheorem{definition}{Definition}

\newcommand{\upset}[1]{#1\!\!\uparrow}
\newcommand{\downset}[1]{#1\!\!\downarrow}
\newcommand{\N}{\mathbb{N}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\SB}{\mathcal{S}}
\newcommand{\Ini}[1]{\textrm{Ini}(#1)}
\newcommand{\Term}[1]{\textrm{Term}(#1)}
\newcommand{\A}{\mathcal{A}}
\newcommand{\powerset}{\mathcal{P}}
\newcommand{\pathway}[1]{\langle#1\rangle}
\newcommand{\R}{\mathbb{R}}

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\green}[1]{{\color{ForestGreen}#1}}
\newcommand{\blue}[1]{{\color{blue}#1}}
\newcommand{\TODO}[1]{\noindent\red{\textbf{TODO:}~#1}}

\begin{document}

\title{Supplementary Information: Quantifying the pathways to life using assembly spaces}
\author{Stuart M. Marshall}
\author{Douglas G. Moore}
\author{Alastair R. G. Murray}
\author{Sara I. Walker}
\author{Leroy Cronin}

\maketitle

\section{Introduction}

This formalism arose as a means of rigorously describing the ``simplest" way of assembling a given object by combining basic building blocks.
With this in mind, we consider a universe of objects together with transitions representing ways in which objects can be combined, mutated or disassembled to yield other objects.
In this setting, the concrete rules or laws relating the transitions with the objects is largely neglected herein, though they are certainly necessary for initially constructing the space for consideration.
We quickly come to the conclusion that a Petri network is appropriate when we consider this kind of process in general, though with specific modifications to allow the types of questions in which we are interested to be well defined.
As such, the fundamental mathematical structure we define and explore herein, referred to as an assembly space, can be described as a sourceless, sinkless Petri net with a chosen basis.
The details of this description will, of course, be made precise below.
With assembly spaces defined, we are then able to define an assembly index for each object in the space, relative to the chosen basis, which characterizes how directly that object can be assembled.
Further, we prove several theorems relating to method of bounding the assembly index for a given object and algorithms for computing or approximating it.

[This is Sara testing if she can git cool like the rest of the peeps in elife, I will remove on my next commit]

\section{Petri Nets}\label{sec:petri-nets}

\subsection{Free Commutative Monoids}\label{subsec:fcm}

The first task we are faced with is how to formally represent the substrates and products of some transition.
We can look to introductory Chemistry for inspiration.
When we consider a process like
\begin{equation*}
    \ce{2H_2} + \ce{O_2} \rightarrow \ce{2H_2O}
\end{equation*}
we are expressing the idea that some chemical process (transition) consumes twice as much $\ce{H_2}$ as $\ce{O_2}$ and yields $\ce{H_2O}$.
The apparent multiplication ($\ce{2H_2}$) and addition ($\ce{2H_2} + \ce{O_2}$) are simply formal constructs use to encode the stoichiometry of the reaction.
The addition here is commutative; we could have written $\ce{O_2} + \ce{2H_2}$, and it would have meant precisely the same thing.

We take a similar approach herein, though mathematically explicit. Let $X$ be a set. This may be, for example, a set of types of molecules $\{\ce{H_2}, \ce{O_2}, \ce{H_2O}\}$ or a set of kinds of building materials and products $\{\textrm{wood}, \textrm{glass pane}, \textrm{nail}, \textrm{picture frame}\}$.
We can then construct formal linear combinations of the elements of $X$ where the coefficients are natural numbers, e.g. $\ce{2H_2} + \ce{O_2}$ or $4\textrm{wood} + \textrm{glass pane} + 8\textrm{nail}$, and only finitely many coefficients are nonzero.

The collection of such formal linear combinations is referred to as the \textit{free commutative monoid} on $X$ over the natural number $\N$, and is typically denoted $\N[X]$.
The set $X$, in this context, is referred to as the set of generators of $\N[X]$.
An element $a \in \N[X]$ can be written as
\begin{equation*}
    a = \sum_{x \in X} a_x x
\end{equation*}
with natural number coefficients $a_x$, only finitely many non-zero.
This construction naturally has a monoidal structure, hence the name.
You can add such linear combinations together, e.g. $\ce{H_2} + (\ce{H_2} + \ce{O_2}) = \ce{2H_2} + \ce{O_2}$, and there exists an identity element $0$ such that $0 + a = a + 0 = a$ for all $a \in \N[X]$.
It is important to stress at this point that the linear combinations do not need to ``make sense'', at least not yet.
Taking the molecular example, $\ce{H_2}$, $\ce{7H_2} + \ce{O_2}$ and $\ce{H_2} + \ce{H_2O} + \ce{O_2}$ are permissible linear combinations and so they are elements of $\N[X]$.
This does not mean that we will, in practice, consider all of them.
Restrictions on which linear combinations are consider will come into the picture when we introduce transitions.

For some of the definitions we will present in coming sections, it will be useful to denote the subset of generators appearing in a given linear combinations. That is for a given element $a \in \N[X]$, the subset generators with non-zero coefficients in $a$, denoted $\pi : \N[X] \rightarrow \powerset{X}$ where
\begin{equation*}
    \displaystyle \sum_{x \in X} a_x x \mapsto \{ x \in X ~|~ a_x \ne 0 \}.
\end{equation*}
To clarify this point, $\pi(\ce{2H2} + \ce{O_2}) = \{\ce{H_2}, \ce{O_2}\}$.

\red{Describe the relationship between the $\pi$ notation and the addition operator since we use both in subsequent sections.}

Before departing from the topic of the free commutative monoid, we want to note that while it may seem that we are being overly formalistic, we gain some power by taking this approach.
Most importantly, we can consider what happens if you ``map'' one set of objects into another.
As a concrete example, you could imagine mapping each of the molecules $\ce{H_2}$, $\ce{O_2}$ and $\ce{H_2O}$ to the number of Hydrogen atoms they contain, i.e. $2$, $0$, and $2$.
\red{I don't like that example very much, but it's good enough for the moment.}
This mapping process can be ``lifted'' into a map on the linear combinations by distributing the map over addition, so that $\ce{H_2} + \ce{O_2}$ maps to $2 + 0 = 2$. More formally, any function between sets $f : X \rightarrow Y$ can be uniquely extended to a function $\N[f] : \N[X] \rightarrow \N[Y]$ such that
\begin{equation*}
    \sum_{x \in X} a_x x \mapsto \sum_{x \in X} a_x f(x).
\end{equation*}
The value of doing this is that it will allow us to coherently define a map between assembly spaces, to be defined, which will then be useful for proving certain theorems regarding bounds on quantities of interest.

\subsection{Petri Nets}

We wish to represent the idea of assembling objects out of ``simpler'' objects, in some sense of the word.
To formalize this idea, we will build off of the concept of a \textit{Petri net}.
In simple terms, a Petri net is a set of sites which we think of as representing types of objects, e.g. molecules, building materials, characters, etc..., and a set of transitions representing the ways in which sites can be combined to yield other sites.
For the sake of formalism, we somewhat elevate the transitions above the sites in that we mostly consider the ``inputs'' and ``outputs'' of transitions, rather the transitions that a given site participates in.
With this in mind, we formally define a Petri net as follows.

\begin{definition}\label{def:petri}
    A \textbf{Petri net} is a $P$ consisting of
    \begin{itemize}
        \item a set of \textit{sites}, $S(P)$
        \item a set of \textit{transitions}, $T(P)$
        \item a pair of maps $\begin{tikzcd}
                T(P)
                \arrow[r, shift left = 1, "\sigma_P"]
                \arrow[r, shift right = 1, "\tau_P", swap]
                &
                \N[S(P)]
        \end{tikzcd}$.
    \end{itemize}
    We call $\sigma_P$ the source map and $\tau_Q$ the target map.
    Where context is unambiguous we may simply write $\sigma$ and $\tau$, but will generally retain the $T(P)$ and $S(P)$ notation.
\end{definition}

The maps $\sigma$ and $\tau$ assign to each transition a formal linear combination of sites.
Picking up our molecular example from the previous section, we can consider the Petri net with sites $S = \{\ce{H_2}, \ce{O_2}, \ce{H_2O}\}$, a single transition $T = \{\alpha\}$, and source and target maps
\begin{equation*}
    \begin{tikzcd}
        \alpha \arrow[r, mapsto, "\sigma"] & 2\ce{H_2} + \ce{O_2}
        \quad\textrm{and}\quad
        \alpha \arrow[r, mapsto, "\tau"] & 2\ce{H_2O}
    \end{tikzcd}
\end{equation*}

There is a natural way of graphically representing a Petri net as a bipartite multigraph:

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (H2) at (0, 1) {$\ce{H_2}$};
                \node [style=site] (O2) at (0, -1) {$\ce{O_2}$};
                \node [style=site] (H2O) at (4, 0) {$\ce{H_2O}$};
                \node [style=transition] (alpha) at (2, 0) {$\alpha$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (H2) to (alpha);
                \draw [style=arrow, bend left=15] (H2) to (alpha);
                \draw [style=arrow] (O2) to (alpha);
                \draw [style=arrow, bend right=15] (alpha) to (H2O);
                \draw [style=arrow, bend left=15] (alpha) to (H2O);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

We read this as ``\textit{transition $\alpha$ can only proceed if at least two $\ce{H_2}$ and one $\ce{O_2}$ are present in the system, and will yield two $\ce{H_2O}$ when it does}''.
In essence, this is a statement about was can possibly happen, not necessarily what will happen; that's more a question of dynamics.

As a more elaborate example, one which we will repeatedly refer to as we expand on the formalism as it exhibits number of interesting features, we imagine a simple ligation type process where in the characters $A$ and $B$ can be combined into strings and $C$ acts something like a \red{``catalyst''} making some transitions possible if and only if $C$ exists.
Consider the Petri $P$ net described by $S(P) = \{ A, B, C, AA, AB, AAAB \}$ and $T(P) = \{ \alpha, \beta, \gamma, \delta \}$ with source and target maps:

\begin{align*}
    \sigma(\alpha) &= 2A           & \tau(\alpha) &= AA         \\
    \sigma(\beta)  &= A + B        & \tau(\beta)  &= AB         \\
    \sigma(\gamma) &= AA + AB + C  & \tau(\gamma) &= AAAB + C   \\
    \sigma(\delta) &= AAAB         & \tau(\delta) &= A + AA + B
\end{align*}

which we graphically represent as

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A) to (tau1);
                \draw [style=arrow, bend left=15] (A) to (tau1);
                \draw [style=arrow, bend right=15] (A) to (tau2);
                \draw [style=arrow] (tau1) to (AA);
                \draw [style=arrow] (B) to (tau2);
                \draw [style=arrow] (tau2) to (AB);
                \draw [style=arrow] (AA) to (tau3);
                \draw [style=arrow] (AB) to (tau3);
                \draw [style=arrow, bend left=30] (C) to (tau3);
                \draw [style=arrow] (tau3) to (AAAB);
                \draw [style=arrow, bend left=30] (tau3) to (C);
                \draw [style=arrow] (AAAB) to (tau4);
                \draw [style=arrow, bend right=30] (tau4) to (A);
                \draw [style=arrow, bend right=30] (tau4) to (AA);
                \draw [style=arrow, bend left=40] (tau4) to (B);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

This particular example exhibits a number of interesting features.
First, there are transitions wherein one of the input sites is also an output site, namely $\gamma$'s dependence on $C$.
Drawing from chemical parlance, this represents a catalytic effect; the transition can only proceed when the site is present\footnote{The addition of another transitition next to $\gamma$ which does not depend on or yield $C$ can model a process that can proceed without $C$ but is accelerated by it's presence.}.
Second, transitions such as $\delta$ can be viewed as a degradation or disassembly process.
This allows reactions to run in both directions.
To elaborate on this point, we can return to the water example and introduce a second transition allowing for the degradation of water back into diatomic Hydrogen and diatomic Oxygen:

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (H2) at (0, 1) {$\ce{H_2}$};
                \node [style=site] (O2) at (0, -1) {$\ce{O_2}$};
                \node [style=site] (H2O) at (4, 0) {$\ce{H_2O}$};
                \node [style=transition] (alpha) at (2, 0) {$\alpha$};
                \node [style=transition] (beta) at (6, 0) {$\beta$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (H2) to (alpha);
                \draw [style=arrow, bend left=15] (H2) to (alpha);
                \draw [style=arrow] (O2) to (alpha);
                \draw [style=arrow, bend right=15] (alpha) to (H2O);
                \draw [style=arrow, bend left=15] (alpha) to (H2O);
                \draw [style=arrow, bend right=15] (H2O) to (beta);
                \draw [style=arrow, bend left=15] (H2O) to (beta);
                \draw [style=arrow, bend right=25] (beta.north) to (H2);
                \draw [style=arrow, bend right=15] (beta.north) to (H2);
                \draw [style=arrow, bend left=15] (beta.south) to (O2);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

An element of $\N[S(P)]$ is called a \textbf{marking} of the Petri net.
Markings play an important role in modeling dynamical systems with Petri nets, representing the state of the system at some time.
For example, the linear combination $5\ce{H_2} + 2\ce{O_2}$ might be thought to represent a system with $5$ molecules of $\ce{H_2}$ and $2$ molecules of $\ce{O_2}$ and no $\ce{H_2O}$.
From this element, we can see that, at least in principle, the transition $\alpha$ might be carried out at most twice\footnote{This is assuming no water dissociates after formation}.
The question of whether, under the action of some dynamical model, one marking can be transformed or processed into another is referred to as the reachability problem and is remarkably difficult.
While we will not delve into dynamical aspects of Petri nets or the to-be-defined assembly spaces in this work, dynamics will likely be a key element of the formal theory going forward.

It is natural, once a mathematical structure has been defined, to ask how two instantiations of that structure may be mapped to one another.
There is, of course, a well defined notion of a map between Petri nets, termed here a \textit{Petri net morphism}, which maps sites and transitions of one Petri net into another in a way that is compatible and consistent with the source and target maps of both Petri nets.
These Petri net morphisms will play a crucial roll in the establishment of bounds for the assembly index in \Cref{subsec:assembly-index-bounds}.

\begin{definition}\label{def:petri-map}
    A \textbf{Petri net morphism} from a Petri net $P$ to a Petri net $Q$, denoted $f : P \rightarrow Q$, consisists of a pair of functions $T(f) : T(P) \rightarrow T(Q)$ and $S(f): S(P) \rightarrow S(Q))$ such that the following diagrams commute:
    \begin{equation*}
        \begin{tikzcd}
            T(P)
            \arrow[r, "\sigma_P"]
            \arrow[d, "T(f)", swap]
            &
            \N[S(P)]
            \arrow[d, "\N \lbrack S(f) \rbrack"]
            \\
            T(Q)
            \arrow[r, "\sigma_Q"]
            &
            \N[S(Q)]
        \end{tikzcd}
        \quad
        \textrm{and}
        \quad
        \begin{tikzcd}
            T(P)
            \arrow[r, "\tau_P"]
            \arrow[d, "T(f)", swap]
            &
            \N[S(P)]
            \arrow[d, "\N \lbrack S(f) \rbrack"]
            \\
            T(Q)
            \arrow[r, "\tau_Q"]
            &
            \N[S(Q)].
        \end{tikzcd}
    \end{equation*}
    When possible, we will abuse notation and write $f(s)$ or $f(t)$ for sites $s \in S(P)$ and transitions $t \in T(P)$, in lieu of the more verbose $S(f)(s)$ and $T(f)(t)$.
\end{definition}

The commutative diagrams in this definition are what we mean by ``compatible and consistent".
They essentially encode the idea that asking for a transition's inputs/outputs and then mapping those inputs/outputs into the new Petri network will give you the same result as first mapping the transition and then asking for the inputs/outputs of the image.
To clarify this point, consider the two Petri nets $P$ and $Q$

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [scale=3] (P) at (12, 0) {$P$};
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};

                \node [scale=3] (Q) at (12, -6) {$Q$};
                \node [style=site] (W) at (-4, -6) {$W$};
                \node [style=site] (X) at (0, -6) {$X$};
                \node [style=site] (Y) at (4, -6) {$Y$};
                \node [style=site] (Z) at (8, -6) {$Z$};
                \node [style=transition] (tau5) at (-2, -6) {$\phi$};
                \node [style=transition] (tau6) at (2, -6) {$\chi$};
                \node [style=transition] (tau7) at (6, -6) {$\psi$};
                \node [style=transition] (tau8) at (10, -6) {$\omega$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A) to (tau1);
                \draw [style=arrow, bend left=15] (A) to (tau1);
                \draw [style=arrow, bend right=15] (A) to (tau2);
                \draw [style=arrow] (tau1) to (AA);
                \draw [style=arrow] (B) to (tau2);
                \draw [style=arrow] (tau2) to (AB);
                \draw [style=arrow] (AA) to (tau3);
                \draw [style=arrow] (AB) to (tau3);
                \draw [style=arrow, bend left=30] (C) to (tau3);
                \draw [style=arrow] (tau3) to (AAAB);
                \draw [style=arrow, bend left=30] (tau3) to (C);
                \draw [style=arrow] (AAAB) to (tau4);
                \draw [style=arrow, bend right=30] (tau4) to (A);
                \draw [style=arrow, bend right=30] (tau4) to (AA);
                \draw [style=arrow, bend left=40] (tau4) to (B);

                \draw [style=arrow, bend left=15] (W) to (tau5);
                \draw [style=arrow, bend right=15] (W) to (tau5);
                \draw [style=arrow, bend left=15] (tau5) to (X);
                \draw [style=arrow, bend right=15] (tau5) to (X);
                \draw [style=arrow, bend left=15] (X) to (tau6);
                \draw [style=arrow, bend right=15] (X) to (tau6);
                \draw [style=arrow] (tau6) to (Y);
                \draw [style=arrow, bend left=20] (Y) to (tau7);
                \draw [style=arrow] (Y) to (tau7);
                \draw [style=arrow, bend right=20] (Y) to (tau7);
                \draw [style=arrow] (tau7) to (Z);
                \draw [style=arrow, bend left=50] (tau7.south) to (Y.south);
                \draw [style=arrow] (Z) to (tau8);
                \draw [style=arrow, bend left=30] (tau8) to (X);
                \draw [style=arrow, bend right=30] (tau8) to (X);
                \draw [style=arrow, bend right=30] (tau8) to (Y);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

We can construct Petri net morphism $f: P \rightarrow Q$ such that
\begin{align*}
    S(f)(\alpha) &= S(f)(\beta) = \chi    & T(f)&(A)    = T(f)(B) = X \\
    S(f)(\gamma) &= \psi                  & T(f)&(AA)   = T(f)(AB) = T(f)(C) = Y \\
    S(f)(\delta) &= \omega                & T(f)&(AAAB) = Z.
\end{align*}

Graphically, this might be depicted as two sets of arrows $T(f)$ and $S(f)$ which map transitions to transitions and sites to sites, respectively colored blue and red.
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}\label{ex:map}
            \begin{pgfonlayer}{nodelayer}
                \node [scale=3] (P) at (12, 0) {$P$};
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};

                \node [scale=3] (Q) at (12, -6) {$Q$};
                \node [style=site] (W) at (-4, -6) {$W$};
                \node [style=site] (X) at (0, -6) {$X$};
                \node [style=site] (Y) at (4, -6) {$Y$};
                \node [style=site] (Z) at (8, -6) {$Z$};
                \node [style=transition] (tau5) at (-2, -6) {$\phi$};
                \node [style=transition] (tau6) at (2, -6) {$\chi$};
                \node [style=transition] (tau7) at (6, -6) {$\psi$};
                \node [style=transition] (tau8) at (10, -6) {$\omega$};

                \node [text=red, scale=2] (s) at (-2, -2.5) {$S(f)$};
                \node [text=blue, scale=2] (t) at (0, -3.5) {$T(f)$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, draw=gray!80, bend right=15] (A) to (tau1);
                \draw [style=arrow, draw=gray!80, bend left=15] (A) to (tau1);
                \draw [style=arrow, draw=gray!80, bend right=15] (A) to (tau2);
                \draw [style=arrow, draw=gray!80] (tau1) to (AA);
                \draw [style=arrow, draw=gray!80] (B) to (tau2);
                \draw [style=arrow, draw=gray!80] (tau2) to (AB);
                \draw [style=arrow, draw=gray!80] (AA) to (tau3);
                \draw [style=arrow, draw=gray!80] (AB) to (tau3);
                \draw [style=arrow, draw=gray!80, bend left=30] (C) to (tau3);
                \draw [style=arrow, draw=gray!80] (tau3) to (AAAB);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau3) to (C);
                \draw [style=arrow, draw=gray!80] (AAAB) to (tau4);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (A);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (AA);
                \draw [style=arrow, draw=gray!80, bend left=40] (tau4) to (B);

                \draw [style=arrow, draw=gray!80, bend left=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend right=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend left=15] (tau5) to (X);
                \draw [style=arrow, draw=gray!80, bend right=15] (tau5) to (X);
                \draw [style=arrow, draw=gray!80, bend left=15] (X) to (tau6);
                \draw [style=arrow, draw=gray!80, bend right=15] (X) to (tau6);
                \draw [style=arrow, draw=gray!80] (tau6) to (Y);
                \draw [style=arrow, draw=gray!80, bend left=20] (Y) to (tau7);
                \draw [style=arrow, draw=gray!80] (Y) to (tau7);
                \draw [style=arrow, draw=gray!80, bend right=20] (Y) to (tau7);
                \draw [style=arrow, draw=gray!80] (tau7) to (Z);
                \draw [style=arrow, draw=gray!80, bend left=50] (tau7.south) to (Y.south);
                \draw [style=arrow, draw=gray!80] (Z) to (tau8);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau8) to (X);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau8) to (X);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau8) to (Y);

                \draw [style=sitemap, bend right=30] (A) to (X);
                \draw [style=sitemap, bend right=30] (B) to (X);
                \draw [style=sitemap, bend right=30] (AA) to (Y);
                \draw [style=sitemap, bend right=30] (AB) to (Y);
                \draw [style=sitemap, bend right=30] (C) to (Y);
                \draw [style=sitemap, bend right=30] (AAAB) to (Z);
                \draw [style=transitionmap, bend right=30] (tau1) to (tau6);
                \draw [style=transitionmap, bend right=30] (tau2) to (tau6);
                \draw [style=transitionmap, bend right=30] (tau3) to (tau7);
                \draw [style=transitionmap, bend right=30] (tau4) to (tau8);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

It is easy to see by inspection that $f$ makes the diagrams in \cref{def:petri-map} commute; for example,
\begin{equation*}
    \sigma_Q(T(f)(\alpha)) = 2X = \N[S(f)](\sigma_P(\alpha)).
\end{equation*}

With an understanding of how to map Petri nets to one another, we can now turn to structural properties and relationships, starting with the concept of a subnet.
In essense, a Petri subnet is simply a Petri net whose sets of sites and transitions are subsets of another net and whose source and target maps are compatible.

\begin{definition}\label{def:subnet}
    Let $Q$ be a Petri net. A \textbf{Petri subnet} $P$ of $Q$ consists of
    \begin{itemize}
        \item a subset of sites of $Q$, denoted $S(P)$
        \item a subset of transitions of $Q$, denoted $T(P)$
        \item a pair of maps $\begin{tikzcd}
                T(P)
                \arrow[r, shift left = 1, "\sigma_P"]
                \arrow[r, shift right = 1, "\tau_P", swap]
                &
                \N[S(P)]
        \end{tikzcd}$.
    \end{itemize}
    such that $\sigma_P = \sigma_Q|_{T(P)}$ and $\tau_P = \tau_Q|_{T(P)}$.
    We denote this relationship as $P \subseteq Q$.
\end{definition}

For example, each of the following nets are Petri subnets of our ligation net:
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A1) at (0, 2) {$A$};
                \node [style=site] (AA1) at (0, 6) {$AA$};
                \node [style=transition] (tau11) at (0, 4) {$\alpha$};

                \node [style=site] (A2) at (4, 2) {$A$};
                \node [style=site] (B2) at (5.6, 2) {$B$};
                \node [style=site] (AA2) at (4, 6) {$AA$};
                \node [style=site] (AB2) at (5.6, 6) {$AB$};
                \node [style=transition] (tau12) at (4, 4) {$\alpha$};
                \node [style=transition] (tau22) at (5.6, 4) {$\beta$};

                \node [style=site] (A3) at (9.6, 0) {$A$};
                \node [style=site] (B3) at (11.2, 0) {$B$};
                \node [style=site] (C3) at (12, 6) {$C$};
                \node [style=site] (AA3) at (9.6, 4) {$AA$};
                \node [style=site] (AB3) at (11.2, 4) {$AB$};
                \node [style=site] (AAAB3) at (10.4, 8) {$AAAB$};
                \node [style=transition] (tau13) at (9.6, 2) {$\alpha$};
                \node [style=transition] (tau23) at (11.2, 2) {$\beta$};
                \node [style=transition] (tau33) at (10.4, 6) {$\gamma$};

                \node [style=site] (AAAB4) at (16.8, 0) {$AAAB$};
                \node [style=site] (A4) at (16, 4) {$A$};
                \node [style=site] (B4) at (17.6, 4) {$B$};
                \node [style=site] (AA4) at (16, 8) {$AA$};
                \node [style=site] (AB4) at (17.6, 8) {$AB$};
                \node [style=transition] (tau44) at (16.8, 2) {$\delta$};
                \node [style=transition] (tau14) at (16, 6) {$\alpha$};
                \node [style=transition] (tau24) at (17.6, 6) {$\beta$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A1) to (tau11);
                \draw [style=arrow, bend left=15] (A1) to (tau11);
                \draw [style=arrow] (tau11) to (AA1);

                \draw [style=arrow, bend right=15] (A2) to (tau12);
                \draw [style=arrow, bend left=15] (A2) to (tau12);
                \draw [style=arrow, bend right=15] (A2) to (tau22);
                \draw [style=arrow] (B2) to (tau22);
                \draw [style=arrow] (tau12) to (AA2);
                \draw [style=arrow] (tau22) to (AB2);

                \draw [style=arrow, bend right=15] (A3) to (tau13);
                \draw [style=arrow, bend left=15] (A3) to (tau13);
                \draw [style=arrow, bend right=15] (A3) to (tau23);
                \draw [style=arrow] (tau13) to (AA3);
                \draw [style=arrow] (B3) to (tau23);
                \draw [style=arrow] (tau23) to (AB3);
                \draw [style=arrow] (AA3) to (tau33);
                \draw [style=arrow] (AB3) to (tau33);
                \draw [style=arrow, bend left=30] (C3) to (tau33);
                \draw [style=arrow] (tau33) to (AAAB3);
                \draw [style=arrow, bend left=30] (tau33) to (C3);

                \draw [style=arrow] (AAAB4) to (tau44);
                \draw [style=arrow] (tau44) to (A4);
                \draw [style=arrow, bend left=50] (tau44) to (AA4);
                \draw [style=arrow] (tau44) to (B4);
                \draw [style=arrow, bend left=15] (A4) to (tau14);
                \draw [style=arrow, bend right=15] (A4) to (tau14);
                \draw [style=arrow, bend right=15] (A4) to (tau24);
                \draw [style=arrow] (B4) to (tau24);
                \draw [style=arrow] (tau14) to (AA4);
                \draw [style=arrow] (tau24) to (AB4);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

Petri net morphisms and subnets interact well.
In particular, for a Petri net morphism $f : P \rightarrow Q$, the image $f(P)$ in $Q$ is a Petri subnet with source and target maps $\sigma_{f(P)}$ and $\tau_{f(P)}$ defined by restriction of $\sigma_Q$ and $\tau_Q$, respectively.
Similarly, the preimage of a Petri subnet of $Q$ is a Petri subnet of $P$ with source and target maps defined by restriction of $\sigma_P$ and $\tau_P$.
Continuing the example map from our ligation model,
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [scale=3] (P) at (12, 0) {$P$};
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};

                \node [scale=3] (Q) at (12, -6) {$Q$};
                \node [style=site] (W) at (-4, -6) {$W$};
                \node [style=site, fill=green!20, draw=red!50] (X) at (0, -6) {$X$};
                \node [style=site, fill=green!20, draw=red!50] (Y) at (4, -6) {$Y$};
                \node [style=site, fill=green!20, draw=red!50] (Z) at (8, -6) {$Z$};
                \node [style=transition] (tau5) at (-2, -6) {$\phi$};
                \node [style=transition, fill=yellow!20, draw=red!50] (tau6) at (2, -6) {$\chi$};
                \node [style=transition, fill=yellow!20, draw=red!50] (tau7) at (6, -6) {$\psi$};
                \node [style=transition, fill=yellow!20, draw=red!50] (tau8) at (10, -6) {$\omega$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, draw=gray!80, bend right=15] (A) to (tau1);
                \draw [style=arrow, draw=gray!80, bend left=15] (A) to (tau1);
                \draw [style=arrow, draw=gray!80, bend right=15] (A) to (tau2);
                \draw [style=arrow, draw=gray!80] (tau1) to (AA);
                \draw [style=arrow, draw=gray!80] (B) to (tau2);
                \draw [style=arrow, draw=gray!80] (tau2) to (AB);
                \draw [style=arrow, draw=gray!80] (AA) to (tau3);
                \draw [style=arrow, draw=gray!80] (AB) to (tau3);
                \draw [style=arrow, draw=gray!80, bend left=30] (C) to (tau3);
                \draw [style=arrow, draw=gray!80] (tau3) to (AAAB);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau3) to (C);
                \draw [style=arrow, draw=gray!80] (AAAB) to (tau4);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (A);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (AA);
                \draw [style=arrow, draw=gray!80, bend left=40] (tau4) to (B);

                \draw [style=arrow, draw=gray!80, bend left=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend right=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend left=15] (tau5) to (X);
                \draw [style=arrow, draw=gray!80, bend right=15] (tau5) to (X);
                \draw [style=arrow, draw=red!50, bend left=15] (X) to (tau6);
                \draw [style=arrow, draw=red!50, bend right=15] (X) to (tau6);
                \draw [style=arrow, draw=red!50] (tau6) to (Y);
                \draw [style=arrow, draw=red!50, bend left=20] (Y) to (tau7);
                \draw [style=arrow, draw=red!50] (Y) to (tau7);
                \draw [style=arrow, draw=red!50, bend right=20] (Y) to (tau7);
                \draw [style=arrow, draw=red!50] (tau7) to (Z);
                \draw [style=arrow, draw=red!50, bend left=50] (tau7.south) to (Y.south);
                \draw [style=arrow, draw=red!50] (Z) to (tau8);
                \draw [style=arrow, draw=red!50, bend left=30] (tau8) to (X);
                \draw [style=arrow, draw=red!50, bend right=30] (tau8) to (X);
                \draw [style=arrow, draw=red!50, bend right=30] (tau8) to (Y);

                \draw [style=sitemap, bend right=30, shorten >=2mm] (A) to (X);
                \draw [style=sitemap, bend right=30, shorten >=2mm] (B) to (X);
                \draw [style=sitemap, bend right=30, shorten >=2mm] (AA) to (Y);
                \draw [style=sitemap, bend right=30, shorten >=2mm] (AB) to (Y);
                \draw [style=sitemap, bend right=30, shorten >=2mm] (C) to (Y);
                \draw [style=sitemap, bend right=30, shorten >=2mm] (AAAB) to (Z);
                \draw [style=transitionmap, bend right=30, shorten >=2mm] (tau1) to (tau6);
                \draw [style=transitionmap, bend right=30, shorten >=2mm] (tau2) to (tau6);
                \draw [style=transitionmap, bend right=30, shorten >=2mm] (tau3) to (tau7);
                \draw [style=transitionmap, bend right=30, shorten >=2mm] (tau4) to (tau8);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
the image, outlined in light red, is a subnet of $Q$.

When it comes to describing systems like the construction of a string, or the building of a molecule, it is often classically unphysical for objects to be created from nothing or destroyed.
For that reason, we need to be able to identify aspects of a Petri net that might represent exactly such a process: sources and sinks.
\begin{definition}\label{def:source-sink}
    Let $P$ be a Petri net and let $t$ be a transition.
    We say that $t$ is a \textbf{source} if $\sigma(t) = 0$, and is a \textbf{sink} if $\tau(t) = 0$.
    $P$ is said to be \textbf{sourceless} if it has no sources.
    Similarly, $P$ is \textbf{sinkless} if it has no sinks.
\end{definition}

Graphically,
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=transition] (source) at (0, 0) {source};
                \node (A) at (2, 1.6) {};
                \node (B) at (2, 0.8) {};
                \node (C) at (2, 0) {};
                \node (D) at (2, -0.8) {};
                \node (E) at (2, -1.6) {};

                \node (F) at (3, 1.6) {};
                \node (G) at (3, 0.8) {};
                \node (H) at (3, 0) {};
                \node (I) at (3, -0.8) {};
                \node (J) at (3, -1.6) {};
                \node [style=transition] (sink) at (5, 0) {sink};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend left=15] (source) to (A);
                \draw [style=arrow, bend left=15] (source) to (B);
                \draw [style=arrow] (source) to (C);
                \draw [style=arrow, bend right=15] (source) to (D);
                \draw [style=arrow, bend right=15] (source) to (E);

                \draw [style=arrow, bend left=15] (F) to (sink);
                \draw [style=arrow, bend left=15] (G) to (sink);
                \draw [style=arrow] (H) to (sink);
                \draw [style=arrow, bend right=15] (I) to (sink);
                \draw [style=arrow, bend right=15] (J) to (sink);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
It follows immediately from the definition of a Petri map that the image of a sourceless Petri net is sourceless, otherwise the $\sigma$-diagram would not commute.
Similarly, the image of a sinkless Petri net is sinkless.

Up to this point, all of the defintions have been fairly standard, at least in terms of concepts.
We now start constructing apparently new ideas that will allow us to formalize the concepts of an assembly space and the assembly index.
The first of these is the idea of a pathway.
Suppose you begin with a set of objects $X$ and wish to consider ways of building some other set of objects.
The simplest approach is to construct a sequence of transitions such that each transition depends only on $X$ or something produced by an earlier transition.
This is what we refer to as a \textit{pathway}.

\begin{definition}\label{def:pathway}
    Let $P$ be a Petri net.
    A \textbf{pathway} in $P$ is a sequence of transitions $\gamma = \pathway{t_n \ldots t_1}$ of length $n \ge 1$ such that for all $1 \le i \le n$
    \begin{equation*}
        \pi(\sigma(t_i)) \subseteq \bigcup_{j=1}^{i-1} \pi(\tau(t_j)).
    \end{equation*}
    We consider the empty sequence of transitions as a pathway and refer to it as the \textbf{zero pathway}.

    \smallskip
    Let $X$ be a subset of sites in $P$. We say that $\gamma$ is \textit{from $X$} if $\pi(\sigma(t_1)) \subseteq X$, and will sometimes say that $\gamma$ starts at $X$.

    \smallskip
    Let $y$ be site, and let $Y$ be a subset of sites.
    We say $\gamma$ is a \textit{to $y$} if $y \in \pi(\tau(t_n))$, and \textit{to $Y$} if $Y \subseteq \pi(\tau(t_n))$, and will sometimes say that $\gamma$ terminates at $y$ or $Y$.

    \smallskip
    We say that the zero pathway is a pathway from any set $X$ to itself and none other.

    \smallskip
    Let $X$ and $Y$ be sets of sites in $P$.
    We say $X$ is \textbf{pathway connected} to $Y$ if there exists a pathway, possibly the zero pathway, from $X$ to $Y$.

    \red{I'd rather express this idea using $\N$ directly instead of using $\pi$}

    \TODO{Break this definition apart}
\end{definition}

Returning to our ligation example, suppose we have the set $X = \{A, B\}$.
There are, in fact, pathways from $X$, counting the zero pathway.
The other three are $\pathway{\alpha}$, $\pathway{\beta}$ and $\pathway{\alpha\beta}$ and $\pathway{\beta\alpha}$.
We depict these pathways below.
The sites in $X$ are outlined in blue, the transitions are filled green, and relevant edges are red.

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site, draw=blue] (A1) at (0, 0.8) {$A$};
                \node [style=site, draw=blue] (B1) at (0, -0.8) {$B$};
                \node [style=site] (C1) at (6, -1.6) {$C$};
                \node [style=site] (AA1) at (4, 0.8) {$AA$};
                \node [style=site] (AB1) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB1) at (8, 0) {$AAAB$};
                \node [style=transition, fill=green!20] (tau11) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau21) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau31) at (6, 0) {$\gamma$};
                \node [style=transition] (tau41) at (10, 0) {$\delta$};
                \node [scale=3] at (5, -3.6) {$\pathway{\alpha}$};

                \node [style=site, draw=blue] (A2) at (14, 0.8) {$A$};
                \node [style=site, draw=blue] (B2) at (14, -0.8) {$B$};
                \node [style=site] (C2) at (20, -1.6) {$C$};
                \node [style=site] (AA2) at (18, 0.8) {$AA$};
                \node [style=site] (AB2) at (18, -0.8) {$AB$};
                \node [style=site] (AAAB2) at (22, 0) {$AAAB$};
                \node [style=transition] (tau12) at (16, 0.8) {$\alpha$};
                \node [style=transition, fill=green!20] (tau22) at (16, -0.8) {$\beta$};
                \node [style=transition] (tau32) at (20, 0) {$\gamma$};
                \node [style=transition] (tau42) at (24, 0) {$\delta$};
                \node [scale=3] at (19, -3.6) {$\pathway{\beta}$};

                \node [style=site, draw=blue] (A3) at (7, -5.8) {$A$};
                \node [style=site, draw=blue] (B3) at (7, -7.4) {$B$};
                \node [style=site] (C3) at (13, -8.2) {$C$};
                \node [style=site] (AA3) at (11, -5.8) {$AA$};
                \node [style=site] (AB3) at (11, -7.4) {$AB$};
                \node [style=site] (AAAB3) at (15, -6.6) {$AAAB$};
                \node [style=transition, fill=green!20] (tau13) at (9, -5.8) {$\alpha$};
                \node [style=transition, fill=green!20] (tau23) at (9, -7.4) {$\beta$};
                \node [style=transition] (tau33) at (13, -6.6) {$\gamma$};
                \node [style=transition] (tau43) at (17, -6.6) {$\delta$};
                \node [scale=3] at (12, -10.2) {$\pathway{\alpha\beta}, \pathway{\beta\alpha}$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, draw=red, bend right=15] (A1) to (tau11);
                \draw [style=arrow, draw=red, bend left=15] (A1) to (tau11);
                \draw [style=arrow, draw=gray!80, bend right=15] (A1) to (tau21);
                \draw [style=arrow, draw=red] (tau11) to (AA1);
                \draw [style=arrow, draw=gray!80] (B1) to (tau21);
                \draw [style=arrow, draw=gray!80] (tau21) to (AB1);
                \draw [style=arrow, draw=gray!80] (AA1) to (tau31);
                \draw [style=arrow, draw=gray!80] (AB1) to (tau31);
                \draw [style=arrow, draw=gray!80, bend left=30] (C1) to (tau31);
                \draw [style=arrow, draw=gray!80] (tau31) to (AAAB1);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau31) to (C1);
                \draw [style=arrow, draw=gray!80] (AAAB1) to (tau41);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau41) to (A1);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau41) to (AA1);
                \draw [style=arrow, draw=gray!80, bend left=40] (tau41) to (B1);

                \draw [style=arrow, draw=gray!80, bend right=15] (A2) to (tau12);
                \draw [style=arrow, draw=gray!80, bend left=15] (A2) to (tau12);
                \draw [style=arrow, draw=red, bend right=15] (A2) to (tau22);
                \draw [style=arrow, draw=gray!80] (tau12) to (AA2);
                \draw [style=arrow, draw=red] (B2) to (tau22);
                \draw [style=arrow, draw=red] (tau22) to (AB2);
                \draw [style=arrow, draw=gray!80] (AA2) to (tau32);
                \draw [style=arrow, draw=gray!80] (AB2) to (tau32);
                \draw [style=arrow, draw=gray!80, bend left=30] (C2) to (tau32);
                \draw [style=arrow, draw=gray!80] (tau32) to (AAAB2);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau32) to (C2);
                \draw [style=arrow, draw=gray!80] (AAAB2) to (tau42);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau42) to (A2);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau42) to (AA2);
                \draw [style=arrow, draw=gray!80, bend left=40] (tau42) to (B2);

                \draw [style=arrow, draw=red, bend right=15] (A3) to (tau13);
                \draw [style=arrow, draw=red, bend left=15] (A3) to (tau13);
                \draw [style=arrow, draw=red, bend right=15] (A3) to (tau23);
                \draw [style=arrow, draw=red] (tau13) to (AA3);
                \draw [style=arrow, draw=red] (B3) to (tau23);
                \draw [style=arrow, draw=red] (tau23) to (AB3);
                \draw [style=arrow, draw=gray!80] (AA3) to (tau33);
                \draw [style=arrow, draw=gray!80] (AB3) to (tau33);
                \draw [style=arrow, draw=gray!80, bend left=30] (C3) to (tau33);
                \draw [style=arrow, draw=gray!80] (tau33) to (AAAB3);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau33) to (C3);
                \draw [style=arrow, draw=gray!80] (AAAB3) to (tau43);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau43) to (A3);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau43) to (AA3);
                \draw [style=arrow, draw=gray!80, bend left=40] (tau43) to (B3);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}

There are a number of points to make with this example.
First, because we did not include $C$ in our starting set, the pathways cannot include $\gamma$ or consequently $\delta$.
Had $X$ included $C$, there would have been at least one pathway from $X$ to any subset of sites.
This is an important point which we will return to shortly.
Second, the graphical representations of both $\pathway{\alpha\beta}$ and $\pathway{\beta\alpha}$ are identical.
This tells us that the order in which $\alpha$ and $\beta$ fire is unimportant as the result is the same.
Finally, in each case the highlighted sites and transitions form a Petri subnet of the larger net.
In fact, given any set of transitions, we can construct a Petri subnet.

\begin{definition}\label{def:site-completion}
    Let $P$ be a Petri net and let $X$ be a set of transitions in $P$.
    The \textbf{site completion} of $X$, denoted $X^*$ is the Petri subnet of $P$ formed such that
    \begin{itemize}
        \item $T(X^*) = X$
        \item $\displaystyle S(X^*) = \bigcup_{t \in X} \pi(\sigma_P(t) + \tau_P(t))$
        \item $\sigma_{X^*} = \sigma_P|_X$ and $\tau_{X^*} = \tau_P|_X$.
    \end{itemize}
\end{definition}

That the site completion is, in fact, a subnet follows immediately from the definition of a subnet.
Furthermore, since we can view a pathway as an ordered set of transitions, we can simply site complete that set to construct a Petri subnet.
There are several questions we could continue with here, such whether every Petri subnet gives rise to a pathway which uses every transition in the subnet.
However, since questions such as this are immaterial to the problem at hand, we will postpone addressing them to subsequent works.

The value of pathways is that they allow us to consider or construct sets of sites which are sufficient to construct every other object in the space.
We refer to such a subset of objects as a \textit{generating set}.

\begin{definition}\label{def:generating-set}
    Let $P$ be a Petri net and let $X$ be a subset of sites of $P$.
    We say that $X$ is a \textbf{generating set} of $P$ if $X$ is pathway connected to every subset of sites $Y$ in $P$.
    The set of all generating sets of $P$ is denoted $\G_P$.
\end{definition}

To clarify this idea, let us consider a few sets of sites from the ligation example.
The first, and least interesting example, is the set of all sites $\{A, B, AA, AB, C, AAAB\}$.
Clearly every site in the space can be created, or at least provided, from this set.
In fact it is trivially true that for any Petri net $P$, the set of all sites in the net is a generating set, i.e. $\G_P \ne \emptyset$.
Furthermore, the empty set of sites is not generating in this case, though it may be in Petri nets with sources (\Cref{def:source-sink}).
The set $\{A,B\}$ is also not a generating set; what sequence of transitions could yield $C$ or $AAAB$?
It is reassuring to know that not all subsets sites are generating, in this case at least.
However, the set $\{A,B,C\}$ is generating, and we can add any other site to this set and the result will be a generating set, e.g. $\{A,B,C,AAAB\}$.
The sets $\{A, AA, AB, C\}$ and $\{AB, AAAB, C\}$ are generating, but upon close inspection we can see that some of the sites within them are not required, e.g. $A$ or $AA$ in the former, and $AB$ in the latter.
This alludes to the idea of a \textit{minimal} generating set; one from which we can remove no sites and the result still be a generating set.
We can formalize this idea by noting that the set of all generating sets of a Petri net is naturally a subset of the powerset of the sites of the net, $\G_P \subseteq \powerset(S(P))$.
As such, $\G_P$ inherits a partial order from $\powerset(S(P))$ (subset inclusion) within which we can determine if a generating set is minimal.
We refer to a minimal generating set with respect to this partial order as a \textit{basis}.

\begin{definition}\label{def:basis}
    Let $P$ be a Petri net and let $X$ be a generating set of $P$, $X \in \G_P$.
    We say that $X$ is a \textbf{basis} if it is minimal with respect to the set inclusion order on $\G_P$.
\end{definition}

The ligation example admits a choice of basis: $\{A,B,C\}$, $\{A,AB,C\}$, $\{AA,AB,C\}$, or $\{C,AAAB\}$.
While it may seem undesirable, on its face, for a Petri net to admit more than one basis, there is little difference between this and being able to choose the basis of a vector space such as $\R^2$.
We can view each choice of basis as representing a minimal set of starting components from which everything of interest can be constructed; similar to deciding which raw materials to use to build a house, or which chemical reagents to use to synthesis a given set of molecules.
Depending on the context, there may be \textit{a priori} reasons to prefer one basis over another.
For example, in the ligation example, the $\{A,B,C\}$ basis seems somehow preferable as it is composed of monomers while all other bases consist of higher-order components.
To make such a distinction, however, requires additional structure, such as a grading\footnote{Here we use the term \textit{grading} loosely. It could, for example, simply mean a map $g : S(P) \rightarrow \N$ for some Petri net $P$ which is consistent in some way with the graph structure of $P$.} on the sites with which to formally distinguish between bases.
The choice of basis will become important in \Cref{sec:assembly-spaces,sec:assembly-index}, but for now we are only interested in the existence and uniqueness of bases and their preservation under Petri net morphisms.
For this reason, we will leave the question of mechanisms for preferentially choosing a basis for a future work.

Whether or not a Petri net admits a basis at all and whether that basis is unique, are both fundamentally important questions.
The establishment of something like the assembly index of an object relies on the specification of a basis.
Fortuitously, every Petri net admits at least one basis, and in certain cases this basis will be unique.
\begin{lemma}\label{lem:bases-exist}
    Every Petri net admits at least one basis.
\end{lemma}
\begin{proof}
    \TODO{Complete (likely to use Zorn's lemma)}
\end{proof}

If we recall, a generating set (and consequently a basis) is a set of sites from which all others can be reached via a pathway.
However, some objects are priviledged in that there are not pathways which terminate at them.
Consider $C$ in the ligation example:
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node (a) at (-4, 0.8) {};
                \node (b) at (-4, -0.8) {};
                \node (c) at (4, 0) {};
                \node [style=site] (C) at (0, -1.6) {$C$};
                \node [style=site] (AA) at (-2, 0.8) {$AA$};
                \node [style=site] (AB) at (-2, -0.8) {$AB$};
                \node [style=site] (AAAB) at (2, 0) {$AAAB$};
                \node [style=transition] (gamma) at (0, 0) {$\gamma$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend left=30] (C) to (gamma);
                \draw [style=arrow, bend left=30] (gamma) to (C);
                \draw [style=arrow] (AA) to (gamma);
                \draw [style=arrow] (AB) to (gamma);
                \draw [style=arrow] (gamma) to (AAAB);
                \draw [style=arrow] (a) to (AA);
                \draw [style=arrow] (b) to (AB);
                \draw [style=arrow] (AAAB) to (c);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
There is no transition that yields $C$ without depending upon it, so $C$ must be included in any generating set or basis.
This is also true of any site that has no incoming edges, for example.
We refer to such a site as these as \textit{initial}, with the concept of \textit{terminal} defined similarly\footnote{We've chosen this terminology to allude to a possible category-theoretic formulation wherein a Petri net is a category with sets of sites as objects and pathways are morphisms. In that context, sets of sites which are initial/terminal this text's formalism are initial/terminal objects in the category.}
\begin{definition}\label{def:initial-final}
    Let $P$ be a Petri net and let $x$ be a site in $P$.
    We say that $x$ is initial in $P$ if the only pathway terminating at $x$ is the zero pathway.
    Similarly, $x$ is final in $P$ if the only pathway starting at $x$ is the zero pathway.
    The set of all initial sites in $P$ is denoted $\Ini{P}$, and the set of all final sites is denoted $\Term{P}$.
\end{definition}
In the ligation example, the only inital site is $C$ and there are no terminal sites.
However, if we were to remove the $\delta$ transition, for example,
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A) to (tau1);
                \draw [style=arrow, bend left=15] (A) to (tau1);
                \draw [style=arrow, bend right=15] (A) to (tau2);
                \draw [style=arrow] (tau1) to (AA);
                \draw [style=arrow] (B) to (tau2);
                \draw [style=arrow] (tau2) to (AB);
                \draw [style=arrow] (AA) to (tau3);
                \draw [style=arrow] (AB) to (tau3);
                \draw [style=arrow, bend left=30] (C) to (tau3);
                \draw [style=arrow] (tau3) to (AAAB);
                \draw [style=arrow, bend left=30] (tau3) to (C);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
then the site $AAAB$ will be terminal and both the initial sites will be $A$, $B$ and $C$.
Sourceless, acyclic Petri nets (nets without loops) will necessarily contain initial sites, and similarly sinkless, acyclic Petri nets will contain terminal sites.

The value of identifying initial sites is that they are necessarily elements of every basis, and if the set of initial sites is a basis in its own right, then it is the only basis.
\begin{lemma}\label{lem:initial-are-basic}
    Let $P$ be a Petri net.
    For any basis $\B$ of $P$, $\Ini{P} \subseteq \B$.
    Furthermore, if $\Ini{P}$ is a basis, then it is a unique basis for $P$.

    \red{I'm not $100\%$ sure of the last point, but I think it's true. We'll find out when we start writing the proof.}

    \green{This is important for internal use. In all of the cases we have dealt with in practice, e.g. molecular assembly as applied, the inital sites form a basis and so there is no ambiguity in the Nat. Comm. paper recently published.}
\end{lemma}
\begin{proof}
    \TODO{Complete}
\end{proof}

Our final concern regaring Petri nets is whether or not bases are are preserved under Petri net morphisms.
To establish that this is the case, we must first show that if two sets of sites are pathway connected, then they will be pathway connected in the image of a Petri net morphism.
\begin{lemma}\label{lem:image-of-pathways}
    Let $f: P \rightarrow Q$ be a Petri net morphism and let $X, Y \subseteq S(P)$ be pathway connected in $P$.
    The images $f(X)$ and $f(Y)$ are pathway connected in $f(P)$.
\end{lemma}
\begin{proof}
    \TODO{Complete}
\end{proof}

As a specific example, we can consider the pathway $\pathway{\alpha\beta}$ in our ligation model, which starts at $\{A, B\}$ and terminates at $\{AA, AB\}$, under the map describe previously.
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [scale=3] (P) at (12, 0) {$P$};
                \node [style=site, fill=blue!20] (A) at (0, 0.8) {$A$};
                \node [style=site, fill=blue!20] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site, fill=red!20] (AA) at (4, 0.8) {$AA$};
                \node [style=site, fill=red!20] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition, fill=green!20] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition, fill=green!20] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};

                \node [scale=3] (Q) at (12, -6) {$Q$};
                \node [style=site] (W) at (-4, -6) {$W$};
                \node [style=site, fill=blue!20] (X) at (0, -6) {$X$};
                \node [style=site, fill=red!20] (Y) at (4, -6) {$Y$};
                \node [style=site] (Z) at (8, -6) {$Z$};
                \node [style=transition] (tau5) at (-2, -6) {$\phi$};
                \node [style=transition, fill=green!20] (tau6) at (2, -6) {$\chi$};
                \node [style=transition] (tau7) at (6, -6) {$\psi$};
                \node [style=transition] (tau8) at (10, -6) {$\omega$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, line width=1.5, draw=red, bend right=15] (A) to (tau1);
                \draw [style=arrow, line width=1.5, draw=red, bend left=15] (A) to (tau1);
                \draw [style=arrow, line width=1.5, draw=red, bend right=15] (A) to (tau2);
                \draw [style=arrow, line width=1.5, draw=red] (tau1) to (AA);
                \draw [style=arrow, line width=1.5, draw=red] (B) to (tau2);
                \draw [style=arrow, line width=1.5, draw=red] (tau2) to (AB);
                \draw [style=arrow, draw=gray!80] (AA) to (tau3);
                \draw [style=arrow, draw=gray!80] (AB) to (tau3);
                \draw [style=arrow, draw=gray!80, bend left=30] (C) to (tau3);
                \draw [style=arrow, draw=gray!80] (tau3) to (AAAB);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau3) to (C);
                \draw [style=arrow, draw=gray!80] (AAAB) to (tau4);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (A);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (AA);
                \draw [style=arrow, draw=gray!80, bend left=40] (tau4) to (B);

                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (A) to (X);
                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (B) to (X);
                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (AA) to (Y);
                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (AB) to (Y);
                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (C) to (Y);
                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (AAAB) to (Z);
                \draw [style=transitionmap, opacity=0.5, bend right=30, shorten >=2mm] (tau1) to (tau6);
                \draw [style=transitionmap, opacity=0.5, bend right=30, shorten >=2mm] (tau2) to (tau6);
                \draw [style=transitionmap, opacity=0.5, bend right=30, shorten >=2mm] (tau3) to (tau7);
                \draw [style=transitionmap, opacity=0.5, bend right=30, shorten >=2mm] (tau4) to (tau8);

                \draw [style=arrow, draw=gray!80, bend left=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend right=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend left=15] (tau5) to (X);
                \draw [style=arrow, draw=gray!80, bend right=15] (tau5) to (X);
                \draw [style=arrow, line width=1.5, draw=red, bend left=15] (X) to (tau6);
                \draw [style=arrow, line width=1.5, draw=red, bend right=15] (X) to (tau6);
                \draw [style=arrow, line width=1.5, draw=red] (tau6) to (Y);
                \draw [style=arrow, draw=gray!80, bend left=20] (Y) to (tau7);
                \draw [style=arrow, draw=gray!80] (Y) to (tau7);
                \draw [style=arrow, draw=gray!80, bend right=20] (Y) to (tau7);
                \draw [style=arrow, draw=gray!80] (tau7) to (Z);
                \draw [style=arrow, draw=gray!80, bend left=50] (tau7.south) to (Y.south);
                \draw [style=arrow, draw=gray!80] (Z) to (tau8);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau8) to (X);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau8) to (X);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau8) to (Y);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
Here the starting set of sites is highlight in blue, the terminal sites in red, the transitions in green and relevant edges are emboldened red.
The images of $\{A,B\}$ and $\{AA,AB\}$ are $\{X\}$ and $\{Y\}$ respectively, and the pathway $\pathway{\alpha\beta}$ is mapped to $\pathway{\chi}$.
As we've proven, any pathway we choose will map to a pathway.
From this fact, it follows that generating sets are preserved.
\begin{lemma}\label{lem:image-of-generating-sets}
    Let $f: P \rightarrow Q$ be a Petri net morphism and let $X$ be a generating set of $P$.
    Then $f(X)$ is a generating set of $f(P)$.
\end{lemma}
\begin{proof}
    \TODO{Complete}
\end{proof}
Identifying a generating set of the ligation model, say $\{A, AB, C\}$, and mapping it forward we see that the result is a generating set of \textit{within the image}:
\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [scale=3] (P) at (12, 0) {$P$};
                \node [style=site, fill=blue!20] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site, fill=blue!20] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site, fill=blue!20] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};

                \node [scale=3] (Q) at (12, -6) {$Q$};
                \node [style=site] (W) at (-4, -6) {$W$};
                \node [style=site, fill=blue!20] (X) at (0, -6) {$X$};
                \node [style=site, fill=blue!20] (Y) at (4, -6) {$Y$};
                \node [style=site] (Z) at (8, -6) {$Z$};
                \node [style=transition] (tau5) at (-2, -6) {$\phi$};
                \node [style=transition] (tau6) at (2, -6) {$\chi$};
                \node [style=transition] (tau7) at (6, -6) {$\psi$};
                \node [style=transition] (tau8) at (10, -6) {$\omega$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, draw=gray!80, bend right=15] (A) to (tau1);
                \draw [style=arrow, draw=gray!80, bend left=15] (A) to (tau1);
                \draw [style=arrow, draw=gray!80, bend right=15] (A) to (tau2);
                \draw [style=arrow, draw=gray!80] (tau1) to (AA);
                \draw [style=arrow, draw=gray!80] (B) to (tau2);
                \draw [style=arrow, draw=gray!80] (tau2) to (AB);
                \draw [style=arrow, draw=gray!80] (AA) to (tau3);
                \draw [style=arrow, draw=gray!80] (AB) to (tau3);
                \draw [style=arrow, draw=gray!80, bend left=30] (C) to (tau3);
                \draw [style=arrow, draw=gray!80] (tau3) to (AAAB);
                \draw [style=arrow, draw=gray!80, bend left=30] (tau3) to (C);
                \draw [style=arrow, draw=gray!80] (AAAB) to (tau4);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (A);
                \draw [style=arrow, draw=gray!80, bend right=30] (tau4) to (AA);
                \draw [style=arrow, draw=gray!80, bend left=40] (tau4) to (B);

                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (A) to (X);
                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (B) to (X);
                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (AA) to (Y);
                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (AB) to (Y);
                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (C) to (Y);
                \draw [style=sitemap, opacity=0.5, bend right=30, shorten >=2mm] (AAAB) to (Z);
                \draw [style=transitionmap, opacity=0.5, bend right=30, shorten >=2mm] (tau1) to (tau6);
                \draw [style=transitionmap, opacity=0.5, bend right=30, shorten >=2mm] (tau2) to (tau6);
                \draw [style=transitionmap, opacity=0.5, bend right=30, shorten >=2mm] (tau3) to (tau7);
                \draw [style=transitionmap, opacity=0.5, bend right=30, shorten >=2mm] (tau4) to (tau8);

                \draw [style=arrow, draw=gray!80, bend left=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend right=15] (W) to (tau5);
                \draw [style=arrow, draw=gray!80, bend left=15] (tau5) to (X);
                \draw [style=arrow, draw=gray!80, bend right=15] (tau5) to (X);
                \draw [style=arrow, draw=red!50, bend left=15] (X) to (tau6);
                \draw [style=arrow, draw=red!50, bend right=15] (X) to (tau6);
                \draw [style=arrow, draw=red!50] (tau6) to (Y);
                \draw [style=arrow, draw=red!50, bend left=20] (Y) to (tau7);
                \draw [style=arrow, draw=red!50] (Y) to (tau7);
                \draw [style=arrow, draw=red!50, bend right=20] (Y) to (tau7);
                \draw [style=arrow, draw=red!50] (tau7) to (Z);
                \draw [style=arrow, draw=red!50, bend left=50] (tau7.south) to (Y.south);
                \draw [style=arrow, draw=red!50] (Z) to (tau8);
                \draw [style=arrow, draw=red!50, bend left=30] (tau8) to (X);
                \draw [style=arrow, draw=red!50, bend right=30] (tau8) to (X);
                \draw [style=arrow, draw=red!50, bend right=30] (tau8) to (Y);

            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
where the generating set $\{A, AB, C\}$ is mapped into the set $\{X, Y\}$, both blue.
The set $\{X, Y\}$ generates the entire image of $P$, whose edges are red.
This is not to say that the only generating sets in the image of a Petri net will be the image of a generating set, for example $\{X\}$ is generating in the image $f(P)$ but its preimage $\{A,B\}$ is not. The property that Petri net morphisms preserve generating sets is fundamentally important for establishing relationships between generator-dependent quantities computed in one Petri net with those in another.

The Petri net formalism, and the extensions presented in this section, can now be leveraged to define the structure of primary interest in this work: assembly spaces.
These spaces can be seen as a subclass of Petri nets which admit specific features useful for describing the fewest number of steps required to construct a given object, the assembly index.
The concepts of subnets and bases will be a key component in the definition of the assembly index, and Petri net morphisms will be indispensible for establishing bounding the value of the assembly index where explict calculation may be infeasible.

\section{Assembly Spaces, Subspaces and Maps}\label{sec:assembly-spaces}

This this work, we are specifically interested in determing if it is possible, given some set of starting objects, to assembly a desired product through some sequence of joining (or disjoining) operations.
If so, how can we assess the minimum number of steps required in the assembly process?
We will refer to the resulting value as the \textit{assembly index} of the object, and ascribe it to the a combination of the structure of the underlying space of assembly processes (\textit{assembly space}), the chosen set of starting objects (\textit{basis}) and the desired product itself.

We draw on the Petri net formalism described and extended in \Cref{sec:petri-nets} to formally define an \textit{assembly space} as a particular variety of Petri net.

\begin{definition}\label{def:assembly-space}
    An \textbf{assembly space} $(A, \B_A)$ is a sourceless and sinkless Petri net with a chosen basis $\B_A$.
    Where possible, we will simply write $A$ to denote an assembly space $(A, \B_A)$.
\end{definition}

The ligation example, reproduced below and used throughout \Cref{sec:petri-nets}, is in fact an assembly space, provided we endow it with a basis.

\begin{center}
    \scalebox{0.6}{
        \begin{tikzpicture}
            \begin{pgfonlayer}{nodelayer}
                \node [style=site] (A) at (0, 0.8) {$A$};
                \node [style=site] (B) at (0, -0.8) {$B$};
                \node [style=site] (C) at (6, -1.6) {$C$};
                \node [style=site] (AA) at (4, 0.8) {$AA$};
                \node [style=site] (AB) at (4, -0.8) {$AB$};
                \node [style=site] (AAAB) at (8, 0) {$AAAB$};
                \node [style=transition] (tau1) at (2, 0.8) {$\alpha$};
                \node [style=transition] (tau2) at (2, -0.8) {$\beta$};
                \node [style=transition] (tau3) at (6, 0) {$\gamma$};
                \node [style=transition] (tau4) at (10, 0) {$\delta$};
            \end{pgfonlayer}
            \begin{pgfonlayer}{edgelayer}
                \draw [style=arrow, bend right=15] (A) to (tau1);
                \draw [style=arrow, bend left=15] (A) to (tau1);
                \draw [style=arrow, bend right=15] (A) to (tau2);
                \draw [style=arrow] (tau1) to (AA);
                \draw [style=arrow] (B) to (tau2);
                \draw [style=arrow] (tau2) to (AB);
                \draw [style=arrow] (AA) to (tau3);
                \draw [style=arrow] (AB) to (tau3);
                \draw [style=arrow, bend left=30] (C) to (tau3);
                \draw [style=arrow] (tau3) to (AAAB);
                \draw [style=arrow, bend left=30] (tau3) to (C);
                \draw [style=arrow] (AAAB) to (tau4);
                \draw [style=arrow, bend right=30] (tau4) to (A);
                \draw [style=arrow, bend right=30] (tau4) to (AA);
                \draw [style=arrow, bend left=40] (tau4) to (B);
            \end{pgfonlayer}
        \end{tikzpicture}
    }
\end{center}
As previously discussed, this net admits four possible bases: $\{A,B,C\}$, $\{A,AB,C\}$, $\{AA,AB,C\}$, or $\{C,AAAB\}$.
In the all of the examples in this sections, we use $\B = \{A,B,C\}$ as the particular basis unless and refer to this this Petri net with $\B$ as the assembly space $\A$.

Assembly spaces, as with Petri nets in general, admit a subnet structure.
In general, a Petri subnet need now have a specified basis; however, we will require this of subnets of assembly spaces.
To distinguish the two concepts, we will refer exclusively to \textit{assembly subspaces} which are little more than Petri subnets with a specified basis.
\begin{definition}\label{def:assembly-subspace}
    Let $A$ and $A'$ be assembly spaces.
    $A'$ is an \textbf{assembly subspace} of $A$ if $A'$ is a Petri subnet of $A$.
    This relationship is denoted $A' \subseteq A$.
\end{definition}

\TODO{Example Figure}

Assembly subspaces whose basis is a subset of the larger space, such as \red{subexample} from the previous example, will be referred to as \textit{rooted}.
\begin{definition}\label{def:rooted}
    Let $(A', \B_{A'})$ be an assembly subspace of $(A, \B_A)$.
    We say that $A'$ is rooted in $A$ if $\B_{A'} \subseteq \B_A$.
\end{definition}
By restricting ourselves to rooted assembly subspaces we tie the particular processes which they represent to the basis of the larger space.
We are particularly interested in constructing assembly subspaces which may produce some specific object.
The subspace relation on the set of all assembly subspaces allows us to partially order the subspaces.
Our ultimate objective is to then constrain the full set of assembly subspaces by requiring they they contain some desired object and determine which subspaces are minimal.
However, if we are do not further restrict this subset of subspace to rooted subspaces, we will inevitably yield highly trivial minimal subspaces.
To elucidate this point, consider the minimal subspaces of $\A$ containing $AB$ with and without the rooting constraint:

\TODO{Example Figure}

Without requiring that the subspaces be rooted, we find that the unenlightening subspace containing only $AB$ as a site and no transitions.
However, given the rooted constraint, we find only one minimal, rooted assembly subspace.
The uniqueness of this minimal rooted subspace is possibly misleading; in general, the minimal subpsaces will not be unique.
This point will be addressed in \Cref{sec:assembly-index}.

As with Petri nets in general, assembly spaces admit structure preserving maps.
In fact, \textit{assembly maps} are just Petri net morphisms between assembly spaces.
\begin{definition}\label{def:assembly-map}
    Let $A$ and $B$ be assembly spaces.
    An \textbf{assembly map} is a Petri net morphism $f: A \rightarrow B$ between assembly spaces.
\end{definition}

As a simple example of this, consider $\A$ and a map to a second space, here referred to as $\C$.

\TODO{Example Figure}

Of particular note here is that $\A$ is certainly a subspace of $\C$, though it is decidedly not rooted in $\C$.
In fact, it will always be the case that the image of an assembly space will be a subspace of the codomain.
\begin{theorem}\label{lem:assembly-image-is-subspace}
    If $f : A \rightarrow B$ is an assembly map between assembly spaces $A$ and $B$, then $f(A)$ is an assembly subspace of $B$.
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}

Exploring the previous example a bit more, by highlighting a rooted subspace of $\A$ we see that the image of that subspace is rooted in the image of $\A$ in $\C$.

\TODO{Example Figure}

This too will always be the case: the image of a rooted subspace will always be rooted in the image.
\begin{theorem}\label{thm:rooted-image}
    Let $A$ be an assembly space, $Z \subseteq A$ a rooted assembly subspace, and $f : A \rightarrow B$ an assembly map.
    Then $(f(Z), S(f)(\B_Z))$ is rooted in $(f(A), S(f)(\B_A)$.

    \TODO{This almost certainly isn't true as defined at present. Assembly maps do not, in general, preserve the basis, $S(f)(\B_A)$ and $S(f)(\B_Z)$ are probably not bases. That said, we might not actually need this. It'll depend on how the proof of \Cref{thm:image-lower-bound} go.}
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}

The interaction of assembly maps and rooted subspaces will allow us to establish bounds on the assembly index, defined in the next section, by considering the image of the space under some map.
These bounds will be useful in situations where determining the assembly index of some object is computationally infeasible.

\section{The Assembly Index}\label{sec:assembly-index}

With the core formalism established, we can now present the concept of the \textit{assembly index}.
As we have seen, within the context of an assembly space, the rooted assembly spaces are partially ordered.
This allows us to inquire about the minimal, rooted assembly subspaces which contain some desirable site.
Each of these subspaces will consist of some, possibly infinite, number of transitions.
We refer to the number of transitions in an assembly space as the \textit{size}.
\begin{definition}\label{def:size}
    Let $A$ be an assembly space.
    The \textbf{size} of $A$ is the cardinality of $T(A)$, denoted $|A|$.
\end{definition}
In the running ligation model, $\A$, the size of the entire space is $|\A| = 4$ as there are four transitions.
Similarly, the minimal, rooted subspace of $\A$ containing $AA$ has size $1$; it contains only one transition, $\alpha$.

We can think of the size of a minimal, rooted subspace containing a given site as counting the number of steps required to assemble that site in a particular way.
However, it is entirely possible that there is no finite way assembly an site.
For example, consider a polymer model which consists of combining monomers $A$ and $B$ into polymers of arbitrary length via pairwise joining operations.
The space of all polymers naturally extends to infinite-length polymers such as $ABABA\ldots$ which require an infinite number of steps to assemble.
While the abstract assembly space may well contain such sites, we wish to restrict ourselves to sites which can be assembled in finitely many steps.
We say that such a site if $\textit{finite}$.
\begin{definition}\label{def:finiteness}
    Let $A$ be an assembly space and let $x$ be a site in $A$.
    We say that $x$ is \textbf{finite} in $A$ there exists a rooted assembly subspace of $A$ which contains $x$ as a site.
\end{definition}
Every site in our running example space, $\A$ is finite.
In the polymer model, only sites representing polymers with finitely many monomers are finite.
If we are presented with such a finite site, we can count the size of every rooted assembly subspaces containing that site.
The least size of these subspaces is the \textit{assembly index} of the site.
\begin{definition}\label{def:index}
    Let $A$ be an assembly space and let $x$ be a finite site in $A$.
    The \textbf{assembly index} of $x$ is the minimum size of all rooted assembly subspaces of $A$ containing $x$, and is denoted $c_A(x)$.
    This can be written $c(x)$ when the relevant assembly space is clear from context.
\end{definition}
In this definition, we do not restrict ourselves specifically to minimal subspaces containing the site $x$.
However, in effect we are.
If we were to consider the size of a non-minimal, rooted subspace some site $x$, we'd always be able to find some minimal, rooted subspace which contains $x$ and has at most the same size.
Since we are looking specifically for the \textit{minimum} size of the subspaces, we need to confirm that such a value even exists.
It does and is in fact unique.
\begin{theorem}\label{def:index-existence-uniqueness}
    Let $A$ be an assembly space and let $x$ be a finite site in $A$.
    Then the assembly index $c_A(x)$ is exists and is unique.
\end{theorem}
\begin{proof}
    \TODO{Complete: likely requires Zorn's lemma, but maybe not...}
\end{proof}

To conclude this section, we present the assembly index of every site of in our ligation model $\A$ together with a minimal, rooted assembly subspace containing that site.

\TODO{Complete Figure}

In the next section, we will consider methods for bounding the assembly index of a site.

\subsection{Computability of and Bounds on the Assembly Index}\label{subsec:assembly-index-bounds}

The first task, when presented with an assembly space, is to determine the assembly index of the sites in the space.
In practical terms, this requires an algorithm for computing the assembly index relative to some space.
In this work, we do not present a general algorithm to do so, though a simple depth-first search of the space may suffice in some cases.
Rather, we first show that the assembly index is in fact computable.
\begin{theorem}\label{thm:computable}
    If $A$ is an assembly space and $x \in S(A)$ is finite, then $c(x)$ is computable.
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}
Of course, just because the value is computable does not mean that it can be computed in reasonable time.
In fact, the general problem is likely to be \texttt{NP}-hard.
In situtations where direct computation is infeasible, we can resort to bounding the value of the assembly index by computing the index in an analogous space.
The first of the bounds we present is an upper bound determined by computing the assembly index in a subspace containing the site.
\begin{theorem}\label{thm:subspace-index-upper-bound}
    Let $(A, \B_A)$ be an assembly space and let $(B, \B_A)$ be a rooted assembly subspace of $A$.
    For every finite $b \in S(B)$, the assembly index of $b$ in $B$ is at least the assembly index of $b$ in $A$.
    That is, $c_B(b) \ge c_A(b)$ for all $b \in S(B)$
\end{theorem}
\begin{proof}
    \TODO{Complete}
\end{proof}

\TODO{Example and Figure}

The second bound we establish is a lower bound determined by mapping the space in question into a space within which computing the assembly index may be simpler.
\begin{theorem}\label{thm:image-lower-bound}
    If $f: A \rightarrow B$ is an assembly map, then $c_{f(A)}(f(x)) \le c_A(x)$ for all finite $x \in S(A)$.
    Furthermore, if $f$ is an embedding, then $c_{f(A)}(f(x)) = c_A(x)$.
\end{theorem}
\begin{proof}
    \TODO{Complete}

    \red{This may suffer from the fact that assembly maps do not preserve bases. If that's the case, we should rip all mention of bases out for a future work.}
\end{proof}

\TODO{Example and Figure}

Our final bound is a special case of the lower bound derived from an assembly map; however, it represents an upper bound.
Given an assembly space with no directed loops, which we refer to as \textbf{acyclic}, we can form an analogous assembly space which has a tree-like structure by duplicating sites that participate in more than one transition.
We refer to such a space as the \textit{split-branch} representation of the space.
\begin{definition}\label{def:split-branch}
    Let $A$ be an acyclic assembly space.
    The \textbf{split-branch} representation of $A$ is $\ldots$.

    \TODO{Doug is leaving this for Stu since it's his jam.}
\end{definition}

There is a natural assembly map from the split-branch representation of a space to the space itself.
As such, \Cref{thm:image-lower-bound} applies, and we can produce an upper bound on the assembly index of sites in the original space.
\begin{corollary}\label{cor:split-branch}
    Let $A$ be a \red{adjectives} assembly space and let $\SB(A)$ be the split-branch representation of $A$, and let $f : \SB(A)$ such that $\ldots$.
    For any site $x$ in $A$, the assembly index $c_A(x) \le \min_{x' \in S(f)^{-1}(x)}c_{\SB(A)}(x')$.
\end{corollary}
\begin{proof}
    \TODO{Stu to complete? Or Doug can with a workable \Cref{def:split-branch}.}
\end{proof}

\TODO{Typeset algorithm}
\red{I need a PDF version of the previous version of the supplement. The formatted algorithm not rendering correctly for me in Word.}

\end{document}